### 2010-07-07 ltx/second_order_framework/second_order_manuscript.tex
model derivation for second order framework

\documentclass[a4paper,10pt]{article}
\usepackage[utf8x]{inputenc}
% *** CITATION PACKAGE ***
\usepackage{cite}
% *** GRAPHICS RELATED PACKAGES ***
\usepackage[pdftex]{graphicx}
% *** MATH PACKAGES ***
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
%Appendix
\usepackage{appendix}
\usepackage{color}

%to delete
\newcommand{\parham}[1]{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}} 
%opening
\title{Non-linear Spatio-temporal modelling using the Integro-Difference Equation}
\author{Parham Aram*,
        Dean~R.~Freestone,Kenneth Scerri, Michael Dewar,\\ Andrew Zammit Mangion and Visakan Kadirkamanathan}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}
\section{State space Representation of the Spatio-temporal Non-linear IDE}
We define the $p$th order non-linear, stochastic, spatio-temporal homogenous IDE as
 \begin{equation}\label{eq:IDEModel}
  \rho_t(\mathbf r)=\int_{\Omega}w(\mathbf r-\mathbf r')f(v_{t-p}(\mathbf r'))d\mathbf r'+\mathbf e_{t-p}(\mathbf r) 
 \end{equation}
where $\rho_t(\mathbf r)$ is a lag polynomial defined as
\begin{equation}\label{eq:LagPolynomial}
 \rho_t(\mathbf r)=(1+\sum_{i=1}^{p}\varphi_i L_i)v_{t}\left(\mathbf r\right)
\end{equation}

where $v_t(\mathbf r)$ denotes the spatial field at time $t$ and at spatial location $\mathbf r$, $f(.)$ is some nonlinear map, and $w(.)$ is the spatial kernel. The disturbance $\mathbf e_{t-p}(\mathbf r)$ is a zero-mean normally distributed noise process, spatially
coloured but temporally independent, with covariance
\begin{equation}
cov\left(e_{t}\left(\mathbf{r}\right),e_{t+\tau}\left(\mathbf{r'}\right)\right)=
\gamma\left(\mathbf{r}-\mathbf{r'}\right)\delta(t-\tau)
\label{eq:FieldCovariance}
\end{equation}
where $\delta(.)$ is the Dirac-delta function.
 The spatial field is observed by
\begin{equation}
    \label{eq:ObservationEquation}
	\mathbf{y}_t =
	\int_{\Omega}{
	    m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)
	d\mathbf{r}'} + 
	\boldsymbol{\varepsilon}_t, 
\end{equation}
where the observation vector $\mathbf{y}_{t}$, compiled at $n_{y}$ spatial locations and at time $t$, is corrupted by an i.i.d normally distributed zero-mean white noise ${\boldsymbol\varepsilon}_t\sim \mathcal{N}\left(\mathbf{0},\boldsymbol\Sigma_{\varepsilon}\right)$ with $\mathbf{\Sigma}_{\boldsymbol\varepsilon}=\sigma_{\varepsilon}^2\mathbf I_{n_y} $. It is assumed here that $\boldsymbol\Sigma_{\varepsilon}$ and $\gamma\left(\mathbf r-\mathbf r' \right)$ are known. The output kernel $m(\mathbf{r}-\mathbf{r}')$ governs the sensor pick-up geometry and is defined by 
\begin{equation}\label{eq:SensorKernel}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)},
\end{equation}
where $\sigma_m$ sets the sensor width. The superscript $\top$ denotes the transpose operator. We proceed by writing the spatio-temporal IDE model in a parameterized state space form which allows the application of standard estimation
techniques. We use a decomposition of the filed and the spatial kernel using a set of Gaussian basis functions
 \begin{equation}
	\label{eq:FieldDecomp} v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t
\end{equation}
\begin{equation}\label{eq:KernelDecomp}
	 w\left(\mathbf{r}-\mathbf{r}'\right) \approx\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right) \boldsymbol{\theta}
\end{equation}
where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ and $\mathbf{\boldsymbol{\psi}}(\mathbf{r})$ are  vectors of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$ and the scaling parameter vector $\boldsymbol\theta$ respectively. Substituting \eqref{eq:LagPolynomial} in \eqref{eq:IDEModel} and using the above decomposition we have
\begin{equation}\label{eq:DecompModel}
\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)(1+\sum_{i=1}^{p}\varphi_i L_i) \mathbf{x}_t=\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right) f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_{t-p})d\mathbf r'\boldsymbol{\theta}+\mathbf e_{t-p}(\mathbf r)
 \end{equation}
multiplying \eqref{eq:DecompModel} by $\boldsymbol \phi(\mathbf r)$ and integrate over the spatial domain $\Omega$ we get
\begin{equation}\label{eq:GammaDecompModel}
 \boldsymbol \Gamma (1+\sum_{i=1}^{p}\varphi_i L_i) \mathbf{x}_t=\int_{\Omega}\boldsymbol \phi(\mathbf r)\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right) f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_{t-p})d\mathbf r'\boldsymbol{\theta}+\int_{\Omega}\boldsymbol\phi(\mathbf r)\mathbf e_{t-p}(\mathbf r)d\mathbf r
\end{equation}
where 
\begin{equation}\label{eq:DefGamma}
	\boldsymbol{\Gamma} \triangleq \int_\Omega {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi} ^{\top}\left(\mathbf{r}\right)d\mathbf{r}} 
\end{equation}
cross-multiplying \eqref{eq:GammaDecompModel} by $\boldsymbol{\Gamma}^{-1}$ gives
\begin{equation}\label{eq:InvGammaDecompModel}
 (1+\sum_{i=1}^{p}\varphi_i L_i) \mathbf{x}_t=\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol \phi(\mathbf r)\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right) f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_{t-p})d\mathbf r'\boldsymbol{\theta}+\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol\phi(\mathbf r)\mathbf e_{t-p}(\mathbf r)d\mathbf r
\end{equation}
This can be further simplified by assuming isotropic structure for the spatial kernel
\begin{equation}
	\boldsymbol{\psi} (\mathbf{r}-\mathbf{r}') = \boldsymbol{\psi} (2c+\mathbf{r}'-\mathbf{r}).
\end{equation}
where $c$ is the center of the kernel. To make the simplification, we first define
\begin{equation}\label{eq:DefPsi}
	\boldsymbol{\Psi}(\mathbf{r}') \triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top} (2c+\mathbf{r}'-\mathbf{r})\textrm{d}\mathbf{r}},
\end{equation}
which is a constant $n_x \times n_{\theta}$ matrix and can be computed analytically. Now substituting \eqref{eq:DefPsi} into \eqref{eq:InvGammaDecompModel} we get
\begin{equation}
(1+\sum_{i=1}^{p}\varphi_i L_i) \mathbf{x}_t = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \boldsymbol{\theta}
+ \boldsymbol{\Gamma}^{-1} \int_\Omega{\boldsymbol{\phi}(\mathbf{r})e_{t-p}(\mathbf{r})\textrm{d}\mathbf{r}}.
\end{equation}
Now we define the state disturbance as
\begin{equation}\label{eq:Wt} 
	\mathbf{e}_{t-p} \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi} ( \mathbf{r} )e_{t-p}( \mathbf{r} )\textrm{d}\mathbf{r}},
\end{equation}
which is a zero mean normally distributed white noise process with covariance
\begin{equation}
	\boldsymbol\Sigma_e =\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 
\end{equation}
the observation equation of the reduced model is found by substituting \eqref{eq:FieldDecomp} into \eqref{eq:ObservationEquation} giving
\begin{equation}\label{ObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where the observation matrix is 
\begin{equation}
	\mathbf{C} = \left[
	\begin{array}{{ccc}} 
		c_{1,1} & \dots & c_{1,L} \\
		\vdots & \ddots & \vdots \\
		c_{N,1} & \dots & c_{N,L} 
	\end{array}
	\right] 
\end{equation}
and 
\begin{equation}
	c_{i,j} = \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}')\textrm{d}\mathbf{r}'. 
\end{equation}
Now we have the final form of the state-space model where
\begin{equation}\label{eq:finalformstatespacemodel}
	(1+\sum_{i=1}^{p}\varphi_i L_i) \mathbf{x}_t = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_{t-p}) \textrm{d}\mathbf{r}' \boldsymbol{\theta}
 +\mathbf{e}_{t-p}
\end{equation}
\begin{equation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t
\end{equation}

Here we consider the case where $p=1$, therefore we have
\begin{equation}
x_t+\varphi_{1} x_{t-1}=\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol \phi(\mathbf r)\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right) f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_{t-1})d\mathbf r'\boldsymbol{\theta}+\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol\phi(\mathbf r)\mathbf e_{t-1}(\mathbf r)d\mathbf r
\end{equation}
\section{Deterministic second order neural field model}
\begin{equation}\label{SpikesToPotential}
	v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'}
\end{equation}
The post-synaptic response kernel $h(t)$ is described by
\begin{equation}\label{SynapticRespKernel}
	h(t) = \eta(t)At\exp{\left(-\zeta t\right)}.
\end{equation}
where $\zeta=\tau^{-1}$ and $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function. $A$ determines the maximal amplitude of the post-synaptic potentials.
Nonlocal interactions between cortical populations are described by	
\begin{equation}\label{RateBasedInteractions}
	g\left( \mathbf{r},t \right) = \int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)d\mathbf{r}'} 
\end{equation}
where $f(.)$ is the firing rate function, $w(.)$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain.
\subsection*{Deriving  Integro-differential Equation}
we can write
\begin{eqnarray}\label{SynapticRespKernelderivation}
 \frac{dh(t)}{dt}&=&A[\delta(t)t\mathrm{exp}(-\zeta t)+\eta(t)\frac{d}{dt}(t\mathrm{exp}(-\zeta t))] \nonumber\\
&=& A[0+\eta(t)[\mathrm{exp}(-\zeta t)-\zeta t\mathrm{exp}(-\zeta t) ]\nonumber \\
&=&A\eta(t)\mathrm{exp}(-\zeta t)-\zeta h(t)
\end{eqnarray}
and taking derivative from \ref{SynapticRespKernelderivation} we have
\begin{eqnarray}\label{SynapticRespKernelSecondderivation}
 \frac{d^2h(t)}{dt^2}&=&A[\delta(t)\mathrm{exp}(-\zeta t)-\zeta\eta(t)\mathrm{exp}(-\zeta t)]-\zeta \frac{dh(t)}{dt} \nonumber\\
&=& A\delta(t)-\zeta A \eta(t)\mathrm{exp}(-\zeta t)-\zeta \frac{dh(t)}{dt} \nonumber \\
&=&A\delta(t)-2\zeta A \eta(t)\mathrm{exp}(-\zeta t)+\zeta^2h(t)
\end{eqnarray}
We define the operator $D$ as $D=\frac{d^2}{dt^2}+2\zeta \frac{d}{dt}+\zeta^2 $, applying $D$ on $h$ and using the result from (\ref{SynapticRespKernelSecondderivation}) it is clear that 
\begin{equation}
 Dh=A\delta(t)
\end{equation}
by applying $D$ on (\ref{SpikesToPotential}) we have
\begin{eqnarray}
 Dv&=&D(h*g) \nonumber \\
&=&(Dh*g) \nonumber\\
&=& A(\delta(t)*g)\nonumber \\
&=&Ag
\end{eqnarray}
therefore we have
\begin{equation}\label{eq:finalcontinouosmodel}
\begin{split}
 \frac{d^2v(\mathbf r,t)}{dt^2}+2\zeta \frac{dv(\mathbf r,t)}{dt}+\zeta^2v(\mathbf r,t)&=Ag\left( \mathbf{r},t \right)\\
&=A\int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)d\mathbf{r}'} \\
\end{split}
\end{equation}
\subsection*{Discritization}
The above second order differential equation can be split into two, coupled first order ODEs.\\
\begin{equation}\label{eq:systemofode}
\begin{cases}
\tilde{v}=\frac{dv}{dt} \\
\frac{d\tilde{v}}{dt}+2\zeta\tilde{v}+\zeta^2v=Ag\left( \mathbf{r},t \right)
\end{cases}
\end{equation}
Now we approximate the two first-order ode by Euler's method
\begin{equation}\label{eq:discritizedsystemofode}
\begin{cases}
v_{t+1}=T_s\tilde{v_t}+v_t\\
\tilde{v}_{t+1}=(1-2\zeta T_s)\tilde{v}_t-\zeta^2T_sv_t+AT_sg_t\left( \mathbf{r}\right)
\end{cases}
\end{equation}
in matrix format we have
\begin{equation}
 \begin{bmatrix}v_{t+1} \\ \tilde{v}_{t+1}\end{bmatrix}= \begin{bmatrix} T_s &1\\ 1-2\zeta T_s & -\zeta^2T_s\end{bmatrix}\begin{bmatrix}v_{t} \\ \tilde{v}_{t}\end{bmatrix}+AT_sg_t\left( \mathbf{r}\right)
\end{equation}
or alternatively we can write (\ref{eq:discritizedsystemofode}) as
\begin{equation}\label{eq:DifferenceModel}
 v_{t+2}-2(1-\zeta T_s)v_{t+1}+(1-2\zeta T_s+\zeta^2T_s^2)v_t=AT_s^2g_t\left( \mathbf{r}\right)
\end{equation}
\subsection*{From Difference equation to state space model}
We define
\begin{equation}\label{eq:tildev}
v_{t+1}-v_t=T_s\tilde{v_t}
\end{equation}
expanding \eqref{eq:DifferenceModel} we get
\begin{equation}\label{eq:rearrangedDifferenceModel}
 v_{t+2}-2v_{t+1}+2\zeta T_S v_{t+1}+v_t-2\zeta T_sv_t+\zeta^2T_s^2v_t=AT_s^2g_t\left( \mathbf{r}\right)
\end{equation}
rearranging \eqref{eq:rearrangedDifferenceModel} we have
\begin{equation}
(v_{t+2}-v_{t+1})-(v_{t+1}-v_t)+2\zeta T_s(v_{t+1}-v_t)+\zeta^2T_s^2v_t=AT_s^2g_t\left( \mathbf{r}\right)
\end{equation}
therefore
\begin{equation}
(v_{t+2}-v_{t+1})-(1-2\zeta T_s)(v_{t+1}-v_t)+\zeta^2T_s^2v_t=AT_s^2g_t\left( \mathbf{r}\right)
\end{equation}
substituting \eqref{eq:tildev} in the above equation and multiplying by $1/T_s$ we get the state space model
\begin{equation}\label{eq:StateSpaceModel}
\begin{cases}
v_{t+1}=T_s\tilde{v_t}+v_t\\
\tilde{v}_{t+1}=(1-2\zeta T_s)\tilde{v}_t-\zeta^2T_sv_t+AT_sg_t\left( \mathbf{r}\right)
\end{cases}
\end{equation}

\section{First order stochastic Neural Field Model}
The model relates the average number of action potentials $g(\mathbf{r},t)$ arriving at time $t \in \mathbb R^{+}$ and at position $\mathbf{r}$ to the local post-synaptic membrane voltage $v(\mathbf{r},t)$. The post-synaptic potentials generated at a neuronal population at location $\mathbf{r}$ by action potentials arriving from all other connected populations at locations $\mathbf{r}'$ is described by 
\begin{equation}
	\label{eq:SpikesToPotential} v\left( {\mathbf{r},t} \right)=v(\mathbf r, 0)h(t)+\int_0^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'} 
\end{equation}
The post-synaptic response kernel $h(t)$ is described by 
\begin{equation}
	\label{eq:SynapticRespKernel} h(t) = \exp{\left(-\zeta t\right)} 
\end{equation}
where $\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant.
Let $W(\mathbf r,t)$ be a spatio-temporal Wiener process. Then, for $\sigma \in \mathbb R^{+}$, the non-local interactions $g\left( {\mathbf{r},t} \right)$ between cortical populations can be divided into a predominantly deterministic part $g'\left( {\mathbf{r},t} \right) $ and  the random component $W(\mathbf r,t)$,
\begin{equation}\label{eq:ActPotComponents}
  g\left( {\mathbf{r},t} \right)dt=g'\left( {\mathbf{r},t} \right)dt+\sigma d W(r,t)
\end{equation}
As a result, the post-synaptic membrane voltage $v(\mathbf r, t)$ is itself a stochastic process. The deterministic part $g'(\mathbf r, t)$ is usually modelled as an integral equation with exogenous deterministic input , $ u(\mathbf r,t)$, of the form \cite{Atay2005}
\begin{equation}
	\label{DeterministicRateBasedInteractions} g'\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)\textrm{d}\mathbf{r}'}+u(\mathbf r,t)
\end{equation}
where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing a cortical sheet or surface. Substituting \eqref{eq:ActPotComponents} in \eqref{eq:SpikesToPotential} we get
\begin{equation}\label{eq:NeuralModelDeterStoch}
v\left( {\mathbf{r},t} \right)=v_0h(t)+\int_0^t {h\left( {t - t'} \right)g'\left( {\mathbf{r},t'} \right)dt'}+ \int_0^t h\left( {t - t'} \right)\sigma d W(r,t')
\end{equation}
Here we assume that $v_0$ is generated by a known distribution $p_0(v)$ and is independent of the the random component $W(\mathbf r,t)$. We stress that $\sigma$ does not depend on the field $v(\mathbf r, t)$
and hence the noise in \eqref{eq:NeuralModelDeterStoch} is strictly additive. It can be shown that the post-synaptic potential $v(\mathbf r, t)$ solves the following initial value problem (see Appendix \ref{app:InitialValueProblemProof})
\begin{equation}
 dv(\mathbf r, t) + \zeta v(\mathbf r, t)dt = g'(\mathbf r, t)dt + dW(\mathbf r, t) \quad t\ge0, v(\mathbf r, 0) = v_0
\end{equation}
applying the Euler-Maruyama method the integro-difference equation (IDE) model takes the form of
\begin{equation}\label{eq:Euler-MaruyamaDiscrit}
 v_{t+T_s}(\mathbf r)- v_{t}(\mathbf r)+ T_s\zeta v_t(\mathbf r)= T_sg'(\mathbf r, t) + (W_{t+T_s}(\mathbf r)-W_t(\mathbf r))
\end{equation}
where $T_s$ is the time step, the increments $W_{t+T_s}(\mathbf r)-W_t(\mathbf r)$ is normally distributed with mean zero and covariance function $T_s\gamma(\mathbf r-\mathbf r')$. We rewrite \eqref{eq:Euler-MaruyamaDiscrit} as
\begin{equation}
	\label{DiscreteTimeModel} 
	v_{t+T_s}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r}'\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\textrm{d}\mathbf{r}'}+T_su_t(\mathbf r)
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
 where $\xi = 1-\zeta T_s$ and $e_t(\mathbf{r})$ is an $i.i.d.$ disturbance such that $e_t(\mathbf{r})\sim\mathcal{GP}(\mathbf 0,T_s\gamma(\mathbf{r}-\mathbf{r}'))$. Here $\mathcal{GP}(\mathbf 0,T_s\gamma(\mathbf{r}-\mathbf{r}'))$ denotes a zero mean spatial Gaussian process with covariance function $T_s\gamma(\mathbf{r}-\mathbf{r}')$. To simplify the notation, the index of the future time sample, $t+T_s$, shall be referred to as $t+1$ throughout the rest of the paper. 
\section{Second order stochastic Neural Field Model}
The model relates the average number of action potentials $g(\mathbf{r},t)$ arriving at time $t \in \mathbb R^{+}$ and at position $\mathbf{r}$ to the local post-synaptic membrane voltage $v(\mathbf{r},t)$. The post-synaptic potentials generated at a neuronal population at location $\mathbf{r}$ by action potentials arriving from all other connected populations at locations $\mathbf{r}'$ is described by 
\begin{equation}
	\label{eq:SecondOrderSpikesToPotential} v\left( {\mathbf{r},t} \right)=v(\mathbf r, 0)(t+1)e^{-\zeta t}+v'(\mathbf r,0)te^{-\zeta t}+\int_0^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'} 
\end{equation}
The post-synaptic response kernel $h(t)$ is described by 
\begin{equation}
	\label{eq:SecondOrderSynapticRespKernel} h(t) = At\exp{\left(-\zeta t\right)} 
\end{equation}
where $\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant.  $A$ determines the maximal amplitude of the post-synaptic potentials.
Let $W(\mathbf r,t)$ be a spatio-temporal Wiener process. Then, for $\sigma \in \mathbb R^{+}$, the non-local interactions $g\left( {\mathbf{r},t} \right)$ between cortical populations can be divided into a predominantly deterministic part $g'\left( {\mathbf{r},t} \right) $ and  the random component $W(\mathbf r,t)$,
\begin{equation}\label{eq:SecondOrderActPotComponents}
  g\left( {\mathbf{r},t} \right)dt=g'\left( {\mathbf{r},t} \right)dt+\sigma d W(r,t)
\end{equation}
As a result, the post-synaptic membrane voltage $v(\mathbf r, t)$ is itself a stochastic process. The deterministic part $g'(\mathbf r, t)$ is usually modelled as an integral equation with exogenous deterministic input , $ u(\mathbf r,t)$, of the form \cite{Atay2005}
\begin{equation}
	\label{DeterministicRateBasedInteractions} g'\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)\textrm{d}\mathbf{r}'}+u(\mathbf r,t)
\end{equation}
where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing a cortical sheet or surface. Substituting \eqref{eq:SecondOrderActPotComponents} in \eqref{eq:SecondOrderSpikesToPotential} we get
\begin{equation}\label{eq:SecondOrderNeuralModelDeterStoch}
v\left( {\mathbf{r},t} \right)=v(\mathbf r, 0)(t+1)e^{-\zeta t}+v'(\mathbf r,0)te^{-\zeta t}+\int_0^t {h\left( {t - t'} \right)g'\left( {\mathbf{r},t'} \right)dt'}+ \int_0^t h\left( {t - t'} \right)\sigma d W(r,t')
\end{equation}
It can be shown that the post-synaptic potential $v(\mathbf r, t)$ solves the following initial value problem \parham{Andrew do you know how?}
\begin{equation*}
\left\lbrace \begin{array}{lc}
dv(\mathbf r,t)=\tilde{v}(\mathbf r,t)d_t & \\
d\tilde{v}(\mathbf r,t)+2\zeta\tilde{v}(\mathbf r,t)dt+\zeta^2v_tdt=&g'(\mathbf r,t)dt+dW(\mathbf r,t)
\end{array}\right.
\end{equation*}
the linear second order SDE can be written in matrix notation as
\begin{equation}
 d\begin{bmatrix} v_t(\mathbf r ,t) \\ \tilde{v}_t(\mathbf r ,t)\end{bmatrix}=\begin{bmatrix}0 & 1 \\ \zeta^2 & 2\zeta \end{bmatrix}\begin{bmatrix} v_t(\mathbf r ,t) \\ \tilde{v}_t(\mathbf r ,t)\end{bmatrix}dt+\begin{bmatrix}0 \\ g'(\mathbf r, t)\end{bmatrix}dt+\begin{bmatrix}0 \\ 1 \end{bmatrix}dW(\mathbf r,t)
\end{equation}

\section{Estimation of the Non-linear Spatio-temporal IDE Model}

\subsection{Unscented Kalman Filter}
\subsection{Unscented RTS Smoother}
\subsection{Parameter Estimation}
\subsection{Computational Complexity}
\section{Example}
\section{Conclusion}

\appendices
\section{}\label{app:InitialValueProblemProof}
%Appendix one text goes here.
Here we show that the post-synaptic potential $v(\mathbf r, t)$ is a solution to the initial value problem 
\begin{equation}
 dv(\mathbf r, t) + \zeta v(\mathbf r, t)dt = g'(\mathbf r, t) + dW(\mathbf r, t) \quad t\ge0, v(\mathbf r, 0) = v_0
\end{equation}
Consider the function $\kappa(v(\mathbf r, t), t) \in C_{1,2}$ defined as
\begin{equation}
\kappa(v(\mathbf r, t), t) = v(\mathbf r, t)e^{\zeta t}
\end{equation}
 applying Ito's formula we get
\begin{equation}\label{eq:kappafunction}
 d\kappa = e^{\zeta t}g'(\mathbf r, t)dt + e^{\zeta t}\sigma dW(\mathbf r, t)
\end{equation}
Integrating \eqref{eq:kappafunction} over $[0, t]$ gives 
\begin{equation}\label{eq:kappafunctionInteration}
 v(\mathbf r, t)e^{\zeta t}-v_0 = \int_0^te^{\zeta t'}g'(\mathbf r, t')dt' +\int_0^te^{\zeta t'}\sigma dW(\mathbf r, t')
\end{equation}
and multiplying throughout by $e^{-\zeta t}$ gives 
\begin{eqnarray}
\begin{split}
 v(\mathbf r, t) &=&v_0e^{-\zeta t}+ \int_0^te^{-\zeta(t-t')}g'(\mathbf r, t')dt' +\int_0^te^{-\zeta (t-t')}\sigma dW(\mathbf r, t')\\
&=&v_0e^{-\zeta t}+ \int_0^t h(t-t')g'(\mathbf r, t')dt' +\int_0^th(t-t')\sigma dW(\mathbf r, t')
\end{split}
\end{eqnarray}
% Now attach the bibliography
\bibliographystyle{plainnat}
\bibliography{STIDE}
\end{document}