### 2010-07-25 ltx/second_order_framework/second_order_manuscript.tex
correction to EM algorithm

\documentclass[a4paper,10pt]{article}
\usepackage[utf8x]{inputenc}
% *** CITATION PACKAGE ***
\usepackage{cite}
% *** GRAPHICS RELATED PACKAGES ***
\usepackage[pdftex]{graphicx}
% *** MATH PACKAGES ***
\usepackage[cmex10]{amsmath}
\usepackage{amssymb}
%Appendix
\usepackage{appendix}
\usepackage{color}

\usepackage{anysize} 								
%<-use this for checking equation length
\marginsize {1.7cm}{1.5cm}{1cm}{1cm} 				%<-use this for
% checking equation length

%to delete
\newcommand{\parham}[1]{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}} 
\newcommand{\dean}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}}

%opening
\title{Nonlinear Spatio-Temporal Modelling using the Integro-Difference
Equation}
\author{Parham Aram*,
        Dean~R.~Freestone,Kenneth Scerri, Michael Dewar,\\ Andrew Zammit Mangion
and Visakan Kadirkamanathan}

\begin{document}

\maketitle

\begin{abstract}
The aim of this paper is to 	
\begin{enumerate}
	\item Derive the temporally second order model
	\item Apply the EM algorithm on test data generated by the model above
	\item Extend the EM algorithm to estimate also the model covariances
	\item Apply method above to measured data
\end{enumerate}

\end{abstract}

\section{Introduction}
\dean{Hi Parham: I have added some headings. Hope they are ok. I was going to
cut and paste the derivation in, but I thought I would let you do it, because it
goes back and forth a little. When we get a structure (doesn't have to be how I
laid it out) it will be easier for me to work on it too.}
\begin{enumerate}
	\item General background and motivation
	\item Aims and hypotheses
	\item Specific background
	\item Point out specific things we are doing and why they are important
	\item Paper layout
\end{enumerate}

\section{Method}
\subsection{Derivation of the Integro-Difference Equation Neural Field Model}

\subsection{Model-Reduction: Derivation of Finite Dimensional State-Space Model}

\subsection{Frequency Analysis: Determining Basis Function Width and Spacing}
Need something about the method of placement of basis functions and sampling of
the field. \dean{Are we going to go for a triangular grid of basis functions?}

\subsection{Determining Connectivity Kernel Width}

\subsection{Estimation: Expectation Maximization Algorithm}

\section{Results}

\subsection{Frequency Analysis}
Confirmation of field sampling by sensors and show spatial frequency constraints
on the approximated field.

\subsection{Connectivity Kernel Width}

\subsection{State and Parameters Estimation}
Need to show the parameter distributions

\section{Discussion and Conclusion}

\section{State-Space Representation of the Spatio-temporal Nonlinear IDE}
The $p^{th}$-order nonlinear, stochastic, spatio-temporal homogenous IDE model
is defined as
 \begin{equation}\label{eq:IDEModel}
  \rho_t(\mathbf r)=\int_{\Omega}w(\mathbf r-\mathbf r')f(v_{t-p}(\mathbf
r'))d\mathbf r'+\mathbf e_{t-p}(\mathbf r). 
 \end{equation}
The quantity $\rho_t(\mathbf r)$ is a lag polynomial defined as
\begin{equation}\label{eq:LagPolynomial}
 \rho_t(\mathbf r)=(1+\sum_{i=1}^{p}\varphi_i L_i)v_{t}\left(\mathbf r\right),
\end{equation}
where $v_t(\mathbf r)$ denotes a spatio-temporal field at time $t$ and at
spatial location $\mathbf r$, $f(\cdot)$ is a nonlinear map, and $w(\cdot)$ is a
spatial mixing kernel. The disturbance, $\mathbf e_{t-p}(\mathbf r)$, is a
zero-mean normally distributed noise process that is spatially
coloured and temporally white, with covariance
\begin{equation}
cov\left(e_{t}\left(\mathbf{r}\right),e_{t+\tau}\left(\mathbf{r'}\right)\right)=
\gamma\left(\mathbf{r}-\mathbf{r'}\right)\delta(t-\tau)
\label{eq:FieldCovariance}
\end{equation}
and $\delta(\cdot)$ is the Dirac-delta function.
Observations from the spatio-temporal field are described by the measurement
function
\begin{equation}
    \label{eq:ObservationEquation}
	\mathbf{y}_t =
	\int_{\Omega}{	   
m\left(\mathbf{r}_{n_y}-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)
	d\mathbf{r}'} + 
	\boldsymbol{\varepsilon}_t, 
\end{equation}
where the observation vector, $\mathbf{y}_{t}$, compiled at $n_{y}$ spatial
locations and at time $t$, is corrupted by an i.i.d normally distributed
zero-mean white noise ${\boldsymbol\varepsilon}_t\sim
\mathcal{N}\left(\mathbf{0},\boldsymbol\Sigma_{\varepsilon}\right)$ with
$\mathbf{\Sigma}_{\boldsymbol\varepsilon}=\sigma_{\varepsilon}^2\mathbf I_{n_y}
$. It is assumed here that $\boldsymbol\Sigma_{\varepsilon}$ and
$\gamma\left(\mathbf r-\mathbf r' \right)$ are known. The output kernel
$m(\mathbf{r}-\mathbf{r}')$ governs the sensor pick-up geometry and is defined
by 
\begin{equation}\label{eq:SensorKernel}
	m\left(\mathbf{r}-\mathbf{r}'\right) =
\exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{
\sigma_m^2}\right)},
\end{equation}
where $\sigma_m$ sets the sensor width. The superscript $\top$ denotes the
transpose operator. We proceed by writing the spatio-temporal IDE model in a
parameterized state space form which allows the application of standard
estimation techniques. We use a decomposition of the field and the spatial
mixing kernel using a set of Gaussian basis functions
 \begin{equation}
	\label{eq:FieldDecomp} v_t\left(\mathbf{r}\right) \approx
\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t
\end{equation}
\begin{equation}\label{eq:KernelDecomp}
	 w\left(\mathbf{r}-\mathbf{r}'\right)
\approx\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right)
\boldsymbol{\theta}
\end{equation}
where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ and
$\mathbf{\boldsymbol{\psi}}(\mathbf{r})$ are  vectors of Gaussian basis
functions that are scaled by the state vector, $\mathbf{x}_t$ and the scaling
parameter vector $\boldsymbol\theta$ respectively. Substituting
\eqref{eq:LagPolynomial} in \eqref{eq:IDEModel} and using the above
decomposition we have
\begin{equation}\label{eq:DecompModel}
\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)(1+\sum_{i=1}^{p}\varphi_i L_i)
\mathbf{x}_t=\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}
'\right) f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right)
\mathbf{x}_{t-p})d\mathbf r'\boldsymbol{\theta}+\mathbf e_{t-p}(\mathbf r)
 \end{equation}
multiplying \eqref{eq:DecompModel} by $\boldsymbol \phi(\mathbf r)$ and
integrate over the spatial domain $\Omega$ we get
\begin{equation}\label{eq:GammaDecompModel}
 \boldsymbol \Gamma (1+\sum_{i=1}^{p}\varphi_i L_i)
\mathbf{x}_t=\int_{\Omega}\boldsymbol \phi(\mathbf
r)\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right)
f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_{t-p})d\mathbf
r'\boldsymbol{\theta}+\int_{\Omega}\boldsymbol\phi(\mathbf r)\mathbf
e_{t-p}(\mathbf r)d\mathbf r
\end{equation}
where 
\begin{equation}\label{eq:DefGamma}
	\boldsymbol{\Gamma} \triangleq \int_\Omega {\boldsymbol{\phi}
\left(\mathbf{r}\right)\boldsymbol{\phi}
^{\top}\left(\mathbf{r}\right)d\mathbf{r}} 
\end{equation}
cross-multiplying \eqref{eq:GammaDecompModel} by $\boldsymbol{\Gamma}^{-1}$
gives
\begin{equation}\label{eq:InvGammaDecompModel}
 (1+\sum_{i=1}^{p}\varphi_i L_i)
\mathbf{x}_t=\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol \phi(\mathbf
r)\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right)
f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_{t-p})d\mathbf
r'\boldsymbol{\theta}+\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol\phi(\mathbf
r)\mathbf e_{t-p}(\mathbf r)d\mathbf r
\end{equation}
This can be further simplified by assuming isotropic structure for the spatial
kernel
\begin{equation}
	\boldsymbol{\psi} (\mathbf{r}-\mathbf{r}') = \boldsymbol{\psi}
(2c+\mathbf{r}'-\mathbf{r}).
\end{equation}
where $c$ is the center of the kernel. \dean{Can't use c as it is used in the
observation matrix.} To make the simplification, we first define
\begin{equation}\label{eq:DefPsi}
	\boldsymbol{\Psi}(\mathbf{r}') \triangleq
T_s\boldsymbol{\Gamma}^{-1}\int_\Omega
{\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top}
(2c+\mathbf{r}'-\mathbf{r})\textrm{d}\mathbf{r}},
\end{equation}
which is a constant $n_x \times n_{\theta}$ matrix and can be computed
analytically. Now substituting \eqref{eq:DefPsi} into
\eqref{eq:InvGammaDecompModel} we get
\begin{equation}
(1+\sum_{i=1}^{p}\varphi_i L_i) \mathbf{x}_t = \int_\Omega
\boldsymbol{\Psi}(\mathbf{r}')
f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}'
\boldsymbol{\theta}
+ \boldsymbol{\Gamma}^{-1}
\int_\Omega{\boldsymbol{\phi}(\mathbf{r})e_{t-p}(\mathbf{r})\textrm{d}\mathbf{r}
}.
\end{equation}
Now we define the state disturbance as
\begin{equation}\label{eq:Wt} 
	\mathbf{e}_{t-p} \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega
{\boldsymbol{\phi} ( \mathbf{r} )e_{t-p}( \mathbf{r} )\textrm{d}\mathbf{r}},
\end{equation}
which is a zero mean normally distributed white noise process with covariance
\begin{equation}
	\boldsymbol\Sigma_e
=\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf
r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf
r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 
\end{equation}
the observation equation of the reduced model is found by substituting
\eqref{eq:FieldDecomp} into \eqref{eq:ObservationEquation} giving
\begin{equation}\label{ObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where the observation matrix is 
\begin{equation}
	\mathbf{C} = \left[
	\begin{array}{{ccc}} 
		c_{1,1} & \dots & c_{1,L} \\
		\vdots & \ddots & \vdots \\
		c_{N,1} & \dots & c_{N,L} 
	\end{array}
	\right] 
\end{equation}
and 
\begin{equation}
	c_{i,j} = \int_{\Omega}m(\mathbf{r}_i -
\mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}')\textrm{d}\mathbf{r}'. 
\end{equation}
Now we have the final form of the state-space model where
\begin{equation}\label{eq:finalformstatespacemodel}
	(1+\sum_{i=1}^{p}\varphi_i L_i) \mathbf{x}_t = \int_\Omega
\boldsymbol{\Psi}(\mathbf{r}')
f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_{t-p}) \textrm{d}\mathbf{r}'
\boldsymbol{\theta}
 +\mathbf{e}_{t-p}
\end{equation}
\begin{equation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t
\end{equation}

Here we consider the case where $p=1$, therefore we have
\begin{equation}
x_t+\varphi_{1} x_{t-1}=\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol
\phi(\mathbf
r)\int_{\Omega}\boldsymbol{\psi}^\top\left(\mathbf{r}-\mathbf{r}'\right)
f(\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_{t-1})d\mathbf
r'\boldsymbol{\theta}+\boldsymbol\Gamma^{-1}\int_{\Omega}\boldsymbol\phi(\mathbf
r)\mathbf e_{t-1}(\mathbf r)d\mathbf r
\end{equation}
\section{Deterministic Second-Order Neural Field Model}
\begin{equation}\label{SpikesToPotential}
	v\left( {\mathbf{r},t} \right) = \int_{ - \infty }^t {h\left( {t - t'}
\right)g\left( {\mathbf{r},t'} \right)dt'}
\end{equation}
The post-synaptic response kernel $h(t)$ is described by
\begin{equation}\label{SynapticRespKernel}
	h(t) = \eta(t)At\exp{\left(-\zeta t\right)}.
\end{equation}
where $\zeta=\tau^{-1}$ and $\tau$ is the synaptic time constant and $\eta(t)$
is the Heaviside step function. $A$ determines the maximal amplitude of the
post-synaptic potentials.
Nonlocal interactions between cortical populations are described by	
\begin{equation}\label{RateBasedInteractions}
	g\left( \mathbf{r},t \right) = \int_\Omega  {w\left(
\mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right)
\right)d\mathbf{r}'} 
\end{equation}
where $f(.)$ is the firing rate function, $w(.)$ is the spatial connectivity
kernel, and $\Omega$ is the spatial domain.
\subsection*{Deriving Integro-differential Equation}
we can write
\begin{eqnarray}\label{SynapticRespKernelderivation}
 \frac{dh(t)}{dt}&=&A[\delta(t)t\mathrm{exp}(-\zeta
t)+\eta(t)\frac{d}{dt}(t\mathrm{exp}(-\zeta t))] \nonumber\\
&=& A[0+\eta(t)[\mathrm{exp}(-\zeta t)-\zeta t\mathrm{exp}(-\zeta t) ]\nonumber
\\
&=&A\eta(t)\mathrm{exp}(-\zeta t)-\zeta h(t)
\end{eqnarray}
and taking derivative from \ref{SynapticRespKernelderivation} we have
\begin{eqnarray}\label{SynapticRespKernelSecondderivation}
 \frac{d^2h(t)}{dt^2}&=&A[\delta(t)\mathrm{exp}(-\zeta
t)-\zeta\eta(t)\mathrm{exp}(-\zeta t)]-\zeta \frac{dh(t)}{dt} \nonumber\\
&=& A\delta(t)-\zeta A \eta(t)\mathrm{exp}(-\zeta t)-\zeta \frac{dh(t)}{dt}
\nonumber \\
&=&A\delta(t)-2\zeta A \eta(t)\mathrm{exp}(-\zeta t)+\zeta^2h(t)
\end{eqnarray}
We define the operator $D$ as $D=\frac{d^2}{dt^2}+2\zeta \frac{d}{dt}+\zeta^2 $,
applying $D$ on $h$ and using the result from
(\ref{SynapticRespKernelSecondderivation}) it is clear that 
\begin{equation}
 Dh=A\delta(t)
\end{equation}
by applying $D$ on (\ref{SpikesToPotential}) we have
\begin{eqnarray}
 Dv&=&D(h*g) \nonumber \\
&=&(Dh*g) \nonumber\\
&=& A(\delta(t)*g)\nonumber \\
&=&Ag
\end{eqnarray}
therefore we have
\begin{equation}\label{eq:finalcontinouosmodel}
\begin{split}
 \frac{d^2v(\mathbf r,t)}{dt^2}+2\zeta \frac{dv(\mathbf
r,t)}{dt}+\zeta^2v(\mathbf r,t)&=Ag\left( \mathbf{r},t \right)\\
&=A\int_\Omega  {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left(
\mathbf{r}',t \right) \right)d\mathbf{r}'} \\
\end{split}
\end{equation}
\subsection*{Discretization}
The above second-order differential equation can be split into two, coupled
first order ODEs.\\
\begin{equation}\label{eq:systemofode}
\begin{cases}
\tilde{v}=\frac{dv}{dt} \\
\frac{d\tilde{v}}{dt}+2\zeta\tilde{v}+\zeta^2v=Ag\left( \mathbf{r},t \right)
\end{cases}
\end{equation}
Now we approximate the two first-order ODE by Euler's method
\begin{equation}\label{eq:discritizedsystemofode}
\begin{cases}
v_{t+1}=T_s\tilde{v_t}+v_t\\
\tilde{v}_{t+1}=(1-2\zeta T_s)\tilde{v}_t-\zeta^2T_sv_t+AT_sg_t\left(
\mathbf{r}\right)
\end{cases}
\end{equation}
in matrix format we have
\begin{equation}
 \begin{bmatrix}v_{t+1} \\ \tilde{v}_{t+1}\end{bmatrix}= \begin{bmatrix} T_s
&1\\ 1-2\zeta T_s & -\zeta^2T_s\end{bmatrix}\begin{bmatrix}v_{t} \\
\tilde{v}_{t}\end{bmatrix}+AT_sg_t\left( \mathbf{r}\right)
\end{equation}
or alternatively we can write (\ref{eq:discritizedsystemofode}) as
\begin{equation}\label{eq:DifferenceModel}
 v_{t+2}-2(1-\zeta T_s)v_{t+1}+(1-2\zeta T_s+\zeta^2T_s^2)v_t=AT_s^2g_t\left(
\mathbf{r}\right)
\end{equation}
\subsection*{From Difference Equation to State-Space Model}
We define
\begin{equation}\label{eq:tildev}
v_{t+1}-v_t=T_s\tilde{v_t}
\end{equation}
expanding \eqref{eq:DifferenceModel} we get
\begin{equation}\label{eq:rearrangedDifferenceModel}
 v_{t+2}-2v_{t+1}+2\zeta T_S v_{t+1}+v_t-2\zeta
T_sv_t+\zeta^2T_s^2v_t=AT_s^2g_t\left( \mathbf{r}\right)
\end{equation}
rearranging \eqref{eq:rearrangedDifferenceModel} we have
\begin{equation}
(v_{t+2}-v_{t+1})-(v_{t+1}-v_t)+2\zeta
T_s(v_{t+1}-v_t)+\zeta^2T_s^2v_t=AT_s^2g_t\left( \mathbf{r}\right)
\end{equation}
therefore
\begin{equation}
(v_{t+2}-v_{t+1})-(1-2\zeta T_s)(v_{t+1}-v_t)+\zeta^2T_s^2v_t=AT_s^2g_t\left(
\mathbf{r}\right)
\end{equation}
substituting \eqref{eq:tildev} in the above equation and multiplying by $1/T_s$
we get the state space model
\begin{equation}\label{eq:StateSpaceModel}
\begin{cases}
v_{t+1}=T_s\tilde{v_t}+v_t\\
\tilde{v}_{t+1}=(1-2\zeta T_s)\tilde{v}_t-\zeta^2T_sv_t+AT_sg_t\left(
\mathbf{r}\right)
\end{cases}
\end{equation}

\section{First order stochastic Neural Field Model}
The model relates the average number of action potentials $g(\mathbf{r},t)$
arriving at time $t \in \mathbb R^{+}$ and at position $\mathbf{r}$ to the local
post-synaptic membrane voltage $v(\mathbf{r},t)$. The post-synaptic potentials
generated at a neuronal population at location $\mathbf{r}$ by action potentials
arriving from all other connected populations at locations $\mathbf{r}'$ is
described by 
\begin{equation}
	\label{eq:SpikesToPotential} v\left( {\mathbf{r},t} \right)=v(\mathbf r,
0)h(t)+\int_0^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'} 
\end{equation}
The post-synaptic response kernel $h(t)$ is described by 
\begin{equation}
	\label{eq:SynapticRespKernel} h(t) = \exp{\left(-\zeta t\right)} 
\end{equation}
where $\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant.
Let $W(\mathbf r,t)$ be a spatio-temporal Wiener process. Then, for $\sigma \in
\mathbb R^{+}$, the non-local interactions $g\left( {\mathbf{r},t} \right)$
between cortical populations can be divided into a predominantly deterministic
part $g'\left( {\mathbf{r},t} \right) $ and  the random component $W(\mathbf
r,t)$,
\begin{equation}\label{eq:ActPotComponents}
  g\left( {\mathbf{r},t} \right)dt=g'\left( {\mathbf{r},t} \right)dt+\sigma d
W(r,t)
\end{equation}
As a result, the post-synaptic membrane voltage $v(\mathbf r, t)$ is itself a
stochastic process. The deterministic part $g'(\mathbf r, t)$ is usually
modelled as an integral equation with exogenous deterministic input , $
u(\mathbf r,t)$, of the form \cite{Atay2005}
\begin{equation}
	\label{DeterministicRateBasedInteractions} g'\left( \mathbf{r},t \right)
= \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left(
\mathbf{r}',t \right) \right)\textrm{d}\mathbf{r}'}+u(\mathbf r,t)
\end{equation}
where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial
connectivity kernel and $\Omega$ is the spatial domain representing a cortical
sheet or surface. Substituting \eqref{eq:ActPotComponents} in
\eqref{eq:SpikesToPotential} we get
\begin{equation}\label{eq:NeuralModelDeterStoch}
v\left( {\mathbf{r},t} \right)=v_0h(t)+\int_0^t {h\left( {t - t'}
\right)g'\left( {\mathbf{r},t'} \right)dt'}+ \int_0^t h\left( {t - t'}
\right)\sigma d W(r,t')
\end{equation}
Here we assume that $v_0$ is generated by a known distribution $p_0(v)$ and is
independent of the the random component $W(\mathbf r,t)$. We stress that
$\sigma$ does not depend on the field $v(\mathbf r, t)$
and hence the noise in \eqref{eq:NeuralModelDeterStoch} is strictly additive. It
can be shown that the post-synaptic potential $v(\mathbf r, t)$ solves the
following initial value problem (see Appendix
\ref{app:InitialValueProblemProof})
\begin{equation}
 dv(\mathbf r, t) + \zeta v(\mathbf r, t)dt = g'(\mathbf r, t)dt + dW(\mathbf r,
t) \quad t\ge0, v(\mathbf r, 0) = v_0
\end{equation}
applying the Euler-Maruyama method the integro-difference equation (IDE) model
takes the form of
\begin{equation}\label{eq:Euler-MaruyamaDiscrit}
 v_{t+T_s}(\mathbf r)- v_{t}(\mathbf r)+ T_s\zeta v_t(\mathbf r)= T_sg'(\mathbf
r, t) + (W_{t+T_s}(\mathbf r)-W_t(\mathbf r))
\end{equation}
where $T_s$ is the time step, the increments $W_{t+T_s}(\mathbf r)-W_t(\mathbf
r)$ is normally distributed with mean zero and covariance function
$T_s\gamma(\mathbf r-\mathbf r')$. We rewrite \eqref{eq:Euler-MaruyamaDiscrit}
as
\begin{equation}
	\label{DiscreteTimeModel} 
	v_{t+T_s}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r}'\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\textrm{d}\mathbf{r}'}+T_su_t(\mathbf r)
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
 where $\xi = 1-\zeta T_s$ and $e_t(\mathbf{r})$ is an $i.i.d.$ disturbance such
that $e_t(\mathbf{r})\sim\mathcal{GP}(\mathbf
0,T_s\gamma(\mathbf{r}-\mathbf{r}'))$. Here $\mathcal{GP}(\mathbf
0,T_s\gamma(\mathbf{r}-\mathbf{r}'))$ denotes a zero mean spatial Gaussian
process with covariance function $T_s\gamma(\mathbf{r}-\mathbf{r}')$. To
simplify the notation, the index of the future time sample, $t+T_s$, shall be
referred to as $t+1$ throughout the rest of the paper. 
\section{Second-Order Stochastic Neural Field Model}
The model relates the average number of action potentials $g(\mathbf{r},t)$
arriving at time $t \in \mathbb R^{+}$ and at position $\mathbf{r}$ to the local
post-synaptic membrane voltage $v(\mathbf{r},t)$. The post-synaptic potentials
generated at a neuronal population at location $\mathbf{r}$ by action potentials
arriving from all other connected populations at locations $\mathbf{r}'$ is
described by 
\begin{equation}
	\label{eq:SecondOrderSpikesToPotential} v\left( {\mathbf{r},t}
\right)=v(\mathbf r, 0)(t+1)e^{-\zeta t}+v'(\mathbf r,0)te^{-\zeta t}+\int_0^t
{h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)dt'} 
\end{equation}
The post-synaptic response kernel $h(t)$ is described by 
\begin{equation}
	\label{eq:SecondOrderSynapticRespKernel} h(t) = At\exp{\left(-\zeta
t\right)} 
\end{equation}
where $\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant.  $A$ determines
the maximal amplitude of the post-synaptic potentials.
Let $W(\mathbf r,t)$ be a spatio-temporal Wiener process. Then, for $\sigma \in
\mathbb R^{+}$, the non-local interactions $g\left( {\mathbf{r},t} \right)$
between cortical populations can be divided into a predominantly deterministic
part $g'\left( {\mathbf{r},t} \right) $ and  the random component $W(\mathbf
r,t)$,
\begin{equation}\label{eq:SecondOrderActPotComponents}
  g\left( {\mathbf{r},t} \right)dt=g'\left( {\mathbf{r},t} \right)dt+\sigma d
W(r,t)
\end{equation}
As a result, the post-synaptic membrane voltage $v(\mathbf r, t)$ is itself a
stochastic process. The deterministic part $g'(\mathbf r, t)$ is usually
modelled as an integral equation with exogenous deterministic input , $
u(\mathbf r,t)$, of the form \cite{Atay2005}
\begin{equation}
	\label{DeterministicRateBasedInteractions} g'\left( \mathbf{r},t \right)
= \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left(
\mathbf{r}',t \right) \right)\textrm{d}\mathbf{r}'}+u(\mathbf r,t)
\end{equation}
where $f(\cdot)$ is the firing rate function, $w(\cdot)$ is the spatial
connectivity kernel and $\Omega$ is the spatial domain representing a cortical
sheet or surface. Substituting \eqref{eq:SecondOrderActPotComponents} in
\eqref{eq:SecondOrderSpikesToPotential} we get
\begin{equation}\label{eq:SecondOrderNeuralModelDeterStoch}
v\left( {\mathbf{r},t} \right)=v(\mathbf r, 0)(t+1)e^{-\zeta t}+v'(\mathbf
r,0)te^{-\zeta t}+\int_0^t {h\left( {t - t'} \right)g'\left( {\mathbf{r},t'}
\right)dt'}+ \int_0^t h\left( {t - t'} \right)\sigma d W(r,t')
\end{equation}
It can be shown that the post-synaptic potential $v(\mathbf r, t)$ solves the
following initial value problem \parham{Andrew do you know how?}
\begin{equation*}\label{eq:SecondOrderInitialValueProblem}
\left\lbrace \begin{array}{lc}
dv(\mathbf r,t)=\tilde{v}(\mathbf r,t)dt & \\
d\tilde{v}(\mathbf r,t)+2\zeta\tilde{v}(\mathbf r,t)dt+\zeta^2v(\mathbf
r,t)dt=&g'(\mathbf r,t)dt+dW(\mathbf r,t)
\end{array}\right.
\end{equation*}
the linear second order SDE can be written in matrix notation as
\begin{equation}\label{eq:SecondOrderSDEMatrix}
 d\begin{bmatrix} v_t(\mathbf r ,t) \\ \tilde{v}_t(\mathbf r
,t)\end{bmatrix}=\begin{bmatrix}0 & 1 \\ -\zeta^2 & -2\zeta
\end{bmatrix}\begin{bmatrix} v_t(\mathbf r ,t) \\ \tilde{v}_t(\mathbf r
,t)\end{bmatrix}dt+\begin{bmatrix}0 \\ g'(\mathbf r,
t)\end{bmatrix}dt+\begin{bmatrix}0 \\ 1 \end{bmatrix}dW(\mathbf r,t)
\end{equation}
applying the Euler-Maruyama method the integro-difference equation (IDE) model
takes the form of
\begin{equation}\label{eq:SecondOrderEulerMaruyamaDis}
 \begin{bmatrix} v_{t+T_s}(\mathbf r)-v_t(\mathbf r) \\
\tilde{v}_{t+T_s}(\mathbf r)-\tilde{v}_t(\mathbf
r)\end{bmatrix}=T_s\begin{bmatrix}0 & 1 \\ -\zeta^2 & -2\zeta
\end{bmatrix}\begin{bmatrix} v_t(\mathbf r)\\ \tilde{v}_t(\mathbf
r)\end{bmatrix}+T_s\begin{bmatrix}0 \\ g'(\mathbf r,
t)\end{bmatrix}+\begin{bmatrix}0 \\ 1 \end{bmatrix}(W_{t+T_s}(\mathbf
r)-W_t(\mathbf r))
\end{equation}
where $T_s$ is the time step, the increments $W_{t+T_s}(\mathbf r)-W_t(\mathbf
r)$ is normally distributed with mean zero and covariance function
$T_s\gamma(\mathbf r-\mathbf r')$. We rewrite
\eqref{eq:SecondOrderEulerMaruyamaDis} as
\begin{equation}
 \begin{bmatrix} v_{t+T_s}(\mathbf r) \\ \tilde{v}_{t+T_s}(\mathbf
r)\end{bmatrix}=\begin{bmatrix}1 & T_s \\ -\zeta^2 T_s & 1-2\zeta T_s
\end{bmatrix}\begin{bmatrix} v_t(\mathbf r)\\ \tilde{v}_t(\mathbf
r)\end{bmatrix}+T_s\begin{bmatrix}0 \\ g'(\mathbf r,
t)\end{bmatrix}+\begin{bmatrix}0 \\ 1 \end{bmatrix}e_t(\mathbf r)
\end{equation}
 where $e_t(\mathbf{r})$ is an $i.i.d.$ disturbance such that
$e_t(\mathbf{r})\sim\mathcal{GP}(\mathbf 0,T_s\gamma(\mathbf{r}-\mathbf{r}'))$.
Here $\mathcal{GP}(\mathbf 0,T_s\gamma(\mathbf{r}-\mathbf{r}'))$ denotes a zero
mean spatial Gaussian process with covariance function
$T_s\gamma(\mathbf{r}-\mathbf{r}')$. To simplify the notation, the index of the
future time sample, $t+T_s$, shall be referred to as $t+1$ throughout the rest
of the paper. 
\section{Estimation of the Non-linear Spatio-temporal IDE Model}

\subsection{Unscented Kalman Filter}
\subsection{Unscented RTS Smoother}
\subsection{Parameter Estimation}
\subsection{Computational Complexity}
\section{Example}
\section{Conclusion}

\appendices
\section{}\label{app:InitialValueProblemProof}
%Appendix one text goes here.
Here we show that the post-synaptic potential $v(\mathbf r, t)$ is a solution to
the initial value problem 
\begin{equation}
 dv(\mathbf r, t) + \zeta v(\mathbf r, t)dt = g'(\mathbf r, t) + dW(\mathbf r,
t) \quad t\ge0, v(\mathbf r, 0) = v_0
\end{equation}
Consider the function $\kappa(v(\mathbf r, t), t) \in C_{1,2}$ defined as
\begin{equation}
\kappa(v(\mathbf r, t), t) = v(\mathbf r, t)e^{\zeta t}
\end{equation}
 applying Ito's formula we get
\begin{equation}\label{eq:kappafunction}
 d\kappa = e^{\zeta t}g'(\mathbf r, t)dt + e^{\zeta t}\sigma dW(\mathbf r, t)
\end{equation}
Integrating \eqref{eq:kappafunction} over $[0, t]$ gives 
\begin{equation}\label{eq:kappafunctionInteration}
 v(\mathbf r, t)e^{\zeta t}-v_0 = \int_0^te^{\zeta t'}g'(\mathbf r, t')dt'
+\int_0^te^{\zeta t'}\sigma dW(\mathbf r, t')
\end{equation}
and multiplying throughout by $e^{-\zeta t}$ gives 
\begin{eqnarray}
\begin{split}
 v(\mathbf r, t) &=&v_0e^{-\zeta t}+ \int_0^te^{-\zeta(t-t')}g'(\mathbf r,
t')dt' +\int_0^te^{-\zeta (t-t')}\sigma dW(\mathbf r, t')\\
&=&v_0e^{-\zeta t}+ \int_0^t h(t-t')g'(\mathbf r, t')dt' +\int_0^th(t-t')\sigma
dW(\mathbf r, t')
\end{split}
\end{eqnarray}
% Now attach the bibliography
\bibliographystyle{plainnat}
\bibliography{STIDE}
\end{document}