### 2010-11-22 ltx/first_order_framework/ResearchHighlights.tex
added tex files for reviewer suggestions and research highlights

%  Created by Dean Freestone on 2010-11-22.
%  Copyright (c) 2010 . All rights reserved.
%
\documentclass[]{article}

\begin{document}

\textbf{Research Highlights}
\begin{itemize}
	\item We derive a model-based, data-driven estimator of continuum neural field parameters.
	\item Systematic field decomposition enables a finite-dimensional state-space model.
	\item We show how to infer intracortical connectivity and synaptic dynamics from data.
	\item Our framework provides a new link between theoretical and experimental neuroscience.
\end{itemize}

\end{document}