### 2011-05-14 ltx/data_paper/DataPaper.tex
minor edits

%
%  untitled
%
%  Created by Dean Freestone on 2010-07-06.
%  Copyright (c) 2010 . All rights reserved.
%
\documentclass[]{article}

% Use utf-8 encoding for foreign characters
\usepackage[utf8]{inputenc}

% Setup for fullpage use
\usepackage{fullpage}

% Uncomment some of the following if you use the features
%
% Running Headers and footers
%\usepackage{fancyhdr}

% Multipart figures
%\usepackage{subfigure}

% More symbols

%\usepackage{latexsym}

% Surround parts of graphics with box
\usepackage{boxedminipage}

% Package for including code in the document
\usepackage{listings}

% If you want to generate a toc for each chapter (use with book)
\usepackage{minitoc}

% This is now the recommended way for checking for PDFLaTeX:
\usepackage{ifpdf}

\usepackage{amsmath,amssymb,amsfonts} % Typical maths resource packages

\newif\ifpdf
\ifx\pdfoutput\undefined
\pdffalse % we are not running PDFLaTeX
\else
\pdfoutput=1 % we are running PDFLaTeX
\pdftrue
\fi

\ifpdf
\usepackage[pdftex]{graphicx}
\else
\usepackage{graphicx}
\fi
\usepackage{color}
\newcommand{\dean}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}} 
\newcommand{\parham}[1]{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}

%\newcommand{\red}{\textcolor{red}}
%\newcommand{\blue}{\textcolor{blue}}
\newcommand{\cyan}{\textcolor{cyan}}

\title{Estimation of Stochastic Wilson and Cowan Neural Field Model using Expectation Maximisation Algorithm}
\author{In no specific order: Parham Aram, Dean R. Freestone, Kenneth Scerri, Michael Dewar,\\
 Andrew Zammit Mangion, Jacob A. Donoghue, Sydney S. Cash, Leigh R. Hochberg,\\
 David B. Grayden and Visakan Kadirkamanathan  }

\date{2010-07-06}

\begin{document}

\ifpdf
\DeclareGraphicsExtensions{.pdf, .jpg, .tif}
\else
\DeclareGraphicsExtensions{.eps, .jpg}
\fi

\maketitle

\begin{abstract}

	\begin{itemize}
		\item Background
		\item Method
		\item Result
		\item Conclusion
	\end{itemize}
\end{abstract}

\section{Introduction}
\begin{itemize}
	\item Understanding of macroscopic neurodynamics
	\item Neural field model
	\item Improvements in electrode technology with great spatiotemporal resolution
\end{itemize}
The aim of this paper is to create a patient-specific neural field model, by estimating parameters of a general mathematical model using data acquired from an epilepsy surgical candidate in a clinical setting.

Instead of extracting features we aim to extract meaningful parameters...

\section{Method}

\subsection{Neural Field Model}
The neural field equations used as the parametric form of the model are based on the influential papers from Wilson and Cowan~\cite{Wilson1973}, and Amari~\cite{Amari1977}. This class of model describes a continuous cortical sheet or surface, relating mean firing rates of pre and post-synaptic neural populations. The neural field, $v\left( {\mathbf{r},t} \right)$, at position $\mathbf{r}$ and time $t$, is the aggregated post-synaptic potentials described by the \dean{(temporal convolution - I'd rather say integral equation	)} integral equation 
\begin{equation}
	\label{SpikesToPotential} v\left( {\mathbf{r},t} \right) = e^{-\zeta t}v\left( {\mathbf{r},0} \right) + \int_{0}^t {h\left( {t - t'} \right)g\left( {\mathbf{r},t'} \right)\textrm{d}t'}, 
\end{equation}   
where $v\left( {\mathbf{r},0} \right)$ is the membrane voltage at the initial time $t = 0$, $h(\cdot)$ is the synaptic response kernel assumed to be first order and of the form 
\begin{equation}
	\label{SynapticRespKernel} h(t) = \eta(t)\exp{\left(-\zeta t\right)}, 
\end{equation}
\cyan{small thing: are we using $\exp(x)$ or $e^x$ to represent exponential?}
$\zeta=\tau^{-1}$, $\tau$ is the synaptic time constant and $\eta(t)$ is the Heaviside step function \parham{(you can keep it here if you wish as it doesn't change the integral equation because of the upper limit. I would remove it)}. $g(\cdot)$ describes the mean firing rate. In practice, the firing rate can rarely be considered as a purely determinstic process because \dean{(can someone write down why?)}. We thus choose to introduce a spatially coloured space-time Wiener process $W(\mathbf{r},t)$ to model the flactuations in the mean firing rate and decompose $g(\cdot)$ as
\begin{equation}\label{DecompositionOfg}
g(\mathbf{r},t)\textrm{d}t = \tilde{g}(\mathbf{r},t)\textrm{d}t + \sigma_W \textrm{d}W(\mathbf{r},t) 
\end{equation}
where $\sigma_W \ge 0$ is a measure of the introduced space-time randomness. For $\sigma_W = 0$ a pure deterministic process is recovered. As a result of the random component, for $\sigma_W > 0$, the membrane voltage $v(\mathbf{r},t)$ is itself a stochastic process. Assuming an infinite propagation velocity for action potentials within the field, the deterministic incoming firing rate is described by a spatial convolution \dean{citation?}
\begin{equation}
	\label{RateBasedInteractions} \tilde{g}\left( \mathbf{r},t \right) = \int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right)f\left( v\left( \mathbf{r}',t \right) \right)\textrm{d}\mathbf{r}'}, 
\end{equation}
where $w(\cdot)$ is the spatial connectivity kernel and $\Omega$ is the spatial domain representing the cortical sheet or surface. The function $f(\cdot)$ relates mean post-synaptic potentials to mean firing rates and follows a sigmoid described by
\begin{equation}
	\label{ActivationFunction} f\left( v\left( \mathbf{r}', t \right) \right) = \frac{1}{1 + \exp \left( \varsigma \left( v_0 - v\left(\mathbf{r}',t\right) \right) \right)}, 
\end{equation}
where $v_0$ and $\varsigma$ describe the firing threshold and the slope of the sigmoid respectively. 
By substituting equation~\ref{DecompositionOfg} into equation~\ref{SpikesToPotential} we obtain the stochastic integral equation model 
\begin{equation}
	\label{StochasticModel} v\left(\mathbf{r},t\right) = e^{-\zeta t}v\left( {\mathbf{r},0} \right) +
	\int_{0}^t h\left(t - t'\right)\tilde{g}\left(\mathbf{r},t'\right) \textrm{d}t'+ \sigma_W\int_{0}^t h\left(t - t'\right)\textrm{d}W\left(\mathbf{r},t'\right).
\end{equation}
 Here we assume that $v\left( {\mathbf{r}},0 \right)$ is independent of the disturbance on the action potentials and assumed to be generated by a known distribution. We stress that $\sigma_W$ does not depend on the field $v({\mathbf{r}},t)$ and hence the noise in equation \ref{StochasticModel} is strictly additive. The general integral-differential equation is given by 
\begin{equation} \label{DifferentialEquationContTime}
 \textrm{d}v(\mathbf{r},t) + \zeta v(\mathbf{r},t)\textrm{d}t = \tilde{g}(\mathbf{r},t)\textrm{d}t + \textrm{d}W(\mathbf{r},t)  ~~~~ t \ge 0, v(\mathbf{r},0) = \cyan{GP?}
\end{equation}
To show that this is indeed the case consider the function $\kappa(v(\mathbf{r},t),t) = v(\mathbf{r},t)e^{\zeta t}$. We note that $\kappa(v(\mathbf{r},t),t)$ is twice differentiable so that we can apply Ito's formula to obtain
\begin{equation}
 \textrm{d}\kappa = e^{\zeta t}g(\mathbf{r},t)\textrm{d}t + e^{\zeta t}\sigma \textrm{d}W(\mathbf{r},t)
\end{equation}
\noindent Integrating over $[0,t]$ and multiplying throughout by $e^{-\zeta t}$ gives the required result.
The neural field equations must be written as a discrete-time finite-dimensional model in order to relate it to patient-specific data. The discrete-time model is found applying the a first-order Euler-Maruyama method on equation \ref{DifferentialEquationContTime} giving
\begin{equation}\label{EulerMethod}
	v\left( \mathbf{r},t+T_s \right)-v\left( \mathbf{r},t \right) + \zeta T_sv\left(\mathbf{r},t \right) = T_s\int_\Omega {w\left( \mathbf{r},\mathbf{r}' \right) f\left( {v\left( \mathbf{r}',t \right)}\right)\textrm{d}\mathbf{r}'} + 	\sigma_W[W\left( \mathbf{r},t+T_s \right)-W\left( \mathbf{r},t \right)],
\end{equation}
where $T_s$ is the time step or sampling period. To simplify the notation, the sample at the the current time step shall be indexed by $t$ and the future time step by $t+1$ for the rest of the paper. Rearranging equation~\ref{EulerMethod} gives the integro-difference equation (IDE) form
\begin{equation}
	\label{NoisyDiscreteTimeModel} 
	v_{t+1}\left(\mathbf{r}\right) = 
	\xi v_t\left(\mathbf{r}\right) + 
	T_s \int_\Omega { 
	    w\left(\mathbf{r},\mathbf{r}'\right)
	    f\left(v_t\left(\mathbf{r}'\right)\right) 
	\textrm{d}\mathbf{r}'} 
	+ e_t\left(\mathbf{r}\right), 
\end{equation}
where $e_t(\mathbf{r}) = \sigma_W[W\left( \mathbf{r},t+T_s \right)-W\left( \mathbf{r},t \right)]$ is the increment of a space-time Wiener process, $i.i.d.$ with zero spatial mean such that $e_t(\mathbf{r})\sim\mathcal{GP}(\mathbf 0,\sigma_d^2\gamma(\mathbf{r}-\mathbf{r}'))$, where $\sigma_d^2=T_s\sigma_W^2$. Here $\mathcal{GP}(\mathbf 0,\sigma_d^2\gamma(\mathbf{r}-\mathbf{r}'))$ denotes a zero mean spatial Gaussian process with covariance function $\gamma(\mathbf{r}-\mathbf{r}')$~\cite{Rasmussen2005}.
To reduce the model to a finite-dimensional system the neural field is approximated by a basis function decomposition where
\begin{equation}
	\label{DefFieldDecomp} v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t 
\end{equation}
and $\mathbf{x}_t$ is a state vector that weights a vector of two-dimension Gaussian basis functions, $\boldsymbol{\phi}(\mathbf{r})$, described by
\begin{equation}\label{eq:FieldBasisFunction}
	\boldsymbol\phi\left(\mathbf{r}-\mathbf{r}'\right) =
\exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_{\phi}^2}\right)}. 
\end{equation}
The parameter $\sigma_{\phi}$ controls the basis function width and is inferred by frequency analysis. Substituting equation~\ref{DefFieldDecomp} into~\ref{NoisyDiscreteTimeModel} gives to approximate model
\begin{equation}
	\label{eq:reduced continuous model}
	\boldsymbol{\phi}^{\top}(\mathbf{r})\mathbf{x}_{t+1}= T_s\int_\Omega{f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t )w(\mathbf{r},\mathbf{r}')\textrm{d}\mathbf{r}'}
	+ \xi\boldsymbol{\phi}^{\top}(\mathbf{r})x_t + e_t(\mathbf{r}). 
\end{equation}
Next we multiply equation~\ref{eq:reduced continuous model} by $\boldsymbol{\phi}(\mathbf{r})$ and integrate over the spatial domain, $\Omega$, to get 
\begin{equation}
	\label{StartofReduction}
 	\int_\Omega {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \textrm{d}\mathbf{r}} \mathbf{x}_{t+1} = T_s \int_\Omega {\boldsymbol{\phi} (\mathbf{r}) \int_\Omega {w(\mathbf{r},\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}') \mathbf{x}_t ) \textrm{d}\mathbf{r}'}\textrm{d}\mathbf{r}} + \xi\int_\Omega{\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\phi}^{\top}(\mathbf{r})\textrm{d}\mathbf{r}} \mathbf{x}_t + \int_\Omega{\boldsymbol{\phi} (\mathbf{r}) e_t(\mathbf{r})\textrm{d}\mathbf{r}}. 
\end{equation}
Now by defining the matrix
\begin{equation}\label{eq:DefGamma}
	\boldsymbol{\Gamma} \triangleq \int_\Omega {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi} ^{\top}\left(\mathbf{r}\right)\textrm{d}\mathbf{r}} 
\end{equation}
enables the pseudo inverse of $\boldsymbol{\phi(\mathbf{r})}$ to be taken, such that the state vector is isolated on the left-hand-side as
\begin{equation}\label{eq:ReducedForm}
	 \mathbf{x}_{t+1} = T_s\boldsymbol{\Gamma}^{-1}
	 \int_\Omega \boldsymbol{\phi}(\mathbf{r}) 
	 \int_\Omega w(\mathbf{r},\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} 
	 + \xi\mathbf{x}_t + \boldsymbol{\Gamma}^{-1} \int_\Omega{\boldsymbol{\phi}(\mathbf{r}) e_t(\mathbf{r})\textrm{d}\mathbf{r}}.
\end{equation}
A property of the basis function decomposition it the preservation of the Gaussianality of the disturbance term. The decomposition disturbance term is defined as
\begin{equation}\label{eq:Wt} 
	\mathbf{e}_t \triangleq \boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} )\textrm{d}\mathbf{r}},
\end{equation}
where $\mathbf{e}_t \sim\mathcal{N}(\mathbf 0,\boldsymbol\Sigma_e)$. The covariance matrix is defined as (see the Supporting Information, S1, for the derivation)
\begin{equation}
	\boldsymbol\Sigma_e \triangleq \sigma_d^2\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \gamma\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top}. 
\end{equation}

To relate the model to the data the electrodes are also modelled. The electrodes have a pick-up range (for LFPs) of approximately 400~$\mu$m. Therefore, we model the electrodes as a spatial convolution of a Gaussian shaped sensor kernel (using a width parameters of 400~$\mu$m) with the neural field. The observation equation is
\begin{equation}
    \label{eq:ObservationEquation}
	\mathbf{y}_t =
	\int_{\Omega}{
	    m\left(\mathbf{r}_n-\mathbf{r}'\right)v_t\left(\mathbf{r}'\right)
	\textrm{d}\mathbf{r}'} + 
	\boldsymbol{\varepsilon}_t, 
\end{equation}
where $\mathbf{r}_n$ defines the location of the sensors in the field, $n=1,...,N$ indexes the sensors and $\boldsymbol{\varepsilon}_t \sim \mathcal{N}\left(0,\boldsymbol{\Sigma}_{\varepsilon}\right)$. $\mathcal{N}\left(0,\boldsymbol{\Sigma}_{\varepsilon}\right)$ denotes the multivariate normal distribution with mean zero and covariance matrix $\boldsymbol{\Sigma}_{\varepsilon}$. The sensor kernel, $m(\mathbf{r}-\mathbf{r}')$, is defined by 
\begin{equation}
	m\left(\mathbf{r}-\mathbf{r}'\right) = \exp{\left(-\frac{(\mathbf{r}-\mathbf{r}')^\top(\mathbf{r}-\mathbf{r}')}{\sigma_m^2}\right)},
\end{equation}
where $\sigma_m$ sets the sensor width. Under the assumption that the basis function decomposition is accurate, the observation equation of the reduced model is found by substituting equation~\ref{DefFieldDecomp} into~\ref{eq:ObservationEquation} giving
\begin{equation}\label{eq:ReducedObservationEquation}
	\mathbf{y}_t = \int_{\Omega}{m\left(\mathbf{r}_n-\mathbf{r}'\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r'}\right) \mathbf{x}_t\textrm{d}\mathbf{r}'} + \boldsymbol{\varepsilon}_t. 
\end{equation}
The observation equation is linear with respect to the state and can be written in the more compact form
\begin{equation}\label{ObservationEquation} 
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{equation}
where the observation matrix is 
\begin{equation}
	\mathbf{C} = \left[
	\begin{array}{{ccc}} 
		c_{1,1} & \dots & c_{1,L} \\
		\vdots & \ddots & \vdots \\
		c_{N,1} & \dots & c_{N,L} 
	\end{array}
	\right] 
\end{equation}
and 
\begin{equation}
	c_{i,j} = \int_{\Omega}m(\mathbf{r}_i - \mathbf{r}')\boldsymbol{\phi}_j(\mathbf{r}')\textrm{d}\mathbf{r}'. 
\end{equation}

\dean{I will do the connectivity kernel decomposition bit after the correlation analysis, because we can use it to say something about heterogeneity and isotropy.}

\subsection{Correlation analysis}\label{subsec:CorrAnal}
Under the assumption that the sensors are not spatially band-limiting the spectral content of the field and the connectivity kernel is homogeneous, the support for the connectivity kernel can be inferred by studying the spatial cross-correlation between consecutive observations. The deterministic component of the spatial mapping between consecutive fields is due to the convolution of the connectivity kernel with the firing rates. Although the relationship between the field and the firing rate is nonlinear, the cross-correlation analysis still extracts meaningful results. This can be seen by considering the three operating regions of the sigmoid. When the mean membrane voltage is below the active region of the sigmoid, the firing rate will be approximately zero and only very small spatial correlations will observed. Therefore, this will not contribute to the results. Within the active region the sigmoid can be approximated by a linear function between consecutive samples, given a sufficiently high sampling rate. Therefore, linear cross-correlation analysis is appropriate. When the sigmoid is in the saturated region, the correlation coefficients will be higher, but not provide a meaningful measure of the kernel weight or gain. Therefore, we can not estimate the gain for the kernel using cross-correlation analysis, but the spatial extent can still be inferred. The gain for the kernel will be inferred using the EM algorithm described below.

Figure~\ref{fig:SpatialCrossCorrelation} shows the average (across time) cross-correlations between consecutive field observations for the different time periods of the seizure progression. By averaging over time, the effect of the disturbance is reduced allowing for the support of the kernel to be seen. The subplots of the figure show that the connectivity kernel exhibits local excitation with surround inhibition. The plots indicate the support for excitatory aspect of the kernel centred at the origin is -2~mm to 2~mm in the $x$ direction and slightly wider in the $y$ direction. Although the kernel is not isotropic, we will approximate it as being isotropic. This is for computation reasons.

The assumption of homogeneity was tested by performing the cross-correlation analysis on 4 equal sized subsections of the field observations. The subsections consisted $6\times6$ electrode grids, such that they overlapped by one electrode row and column. The overlap was required so the subsections where large enough for the support of the kernel to be seen in the cross-correlation analysis. The results of the analysis can be seen in Figure~\ref{fig:HomogeneityTest}. The results show that each subsection of the field has similar cross-correlations, with each subplot showing local excitation with surround inhibition. From this analysis, we will approximate the connectivity kernel as being homogeneous.

\subsection{Parametrisation of Connectivity Kernel}
The connectivity kernel can be decomposed as 
\begin{equation}\label{eq:DefKernelDecomp}
	 w\left(\mathbf{r},\mathbf{r}'\right) \approx \boldsymbol{\psi}^\top\left(\mathbf{r},\mathbf{r}'\right) \boldsymbol{\theta},
\end{equation}
where $\boldsymbol{\psi}(\mathbf{r},\mathbf{r}')$ is a vector of Gaussian basis functions and $\boldsymbol{\theta}$ is a vector of scaling parameters. By assuming a Gaussian isotropic connectivity structure, the kernel basis functions can be written as $\boldsymbol{\psi}(\mathbf{r}-\mathbf{r}')$. We will assume that we know the parametric form of the connectivity basis functions from the correlation analysis, where the scaling parameters $\boldsymbol{\theta}$ are unknown.
Substituting equation~\ref{eq:DefKernelDecomp} into equation\ref{eq:ReducedForm} gives
\begin{equation}\label{eq:KernelDecompForm}
	 \mathbf{x}_{t+1} = T_s\boldsymbol{\Gamma}^{-1}
	 \int_\Omega \boldsymbol{\phi}(\mathbf{r}) 
	 \int_\Omega \boldsymbol{\psi}^\top\left(\mathbf{r},\mathbf{r}'\right) \boldsymbol{\theta} f(\boldsymbol{\phi}^{\top} (\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} 
	 + \xi\mathbf{x}_t + \mathbf{e}_t(\mathbf{r}).
\end{equation}
The connectivity kernel weights can be shifted outside the integral giving
\begin{equation}
	 \mathbf{x}_{t+1} = T_s\boldsymbol{\Gamma}^{-1}
	 \int_\Omega \boldsymbol{\phi}(\mathbf{r}) 
	 \int_\Omega \boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} \boldsymbol{\theta}  
	 + \xi\mathbf{x}_t + \mathbf{e}_t(\mathbf{r}).
\end{equation}
Now taking the basis function vector inside the inner integrand gives
\begin{equation}\label{eq:ReucedFormWithKernelDecomp}
	 \mathbf{x}_{t+1} = T_s\boldsymbol{\Gamma}^{-1}
	 \int_\Omega \int_\Omega \boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top} (\mathbf{r}-\mathbf{r}')f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \textrm{d}\mathbf{r} \boldsymbol{\theta}  
	 + \xi\mathbf{x}_t + \mathbf{e}_t(\mathbf{r}).
\end{equation}
This equation can be simplified by exploiting the isotropy of the basis functions used to represent the kernel, and hence,
\begin{equation}
	\psi_i (\mathbf{r}-\mathbf{r}') = \psi_i (2c_i+\mathbf{r}'-\mathbf{r}).
\end{equation}
where $c_i$ is the center of the $i^{th}$ basis function of the kernel decomposition. To make the simplification, we define
\begin{equation}\label{eq:DefPsi}
	\left[ \boldsymbol\Psi(\mathbf{r}')\right]_{:i}  \triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\psi_i (2c_i+\mathbf{r}'-\mathbf{r})\textrm{d}\mathbf{r}},
\end{equation}
where $\left[ \boldsymbol\Psi(\mathbf{r}')\right]_{:i}$ denotes the $i^{th}$column of $\boldsymbol{\Psi}(\mathbf{r}')$ which is a constant $L \times n_{\theta}$ matrix that is defined analytically, where $L$ is the number of basis functions (and states) and $n_{\theta}$ is the number of connectivity kernel basis functions. Substituting equation~\ref{eq:DefPsi} into~\ref{eq:ReucedFormWithKernelDecomp} gives
\begin{equation}
	\mathbf{x}_{t+1} = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \textrm{d}\mathbf{r}' \boldsymbol{\theta} + \xi\mathbf{x}_t 
+ \mathbf{e}_t(\mathbf{r}).
\end{equation}
This substitution exploits the isotropy of the connectivity kernel basis functions, allowing us to swap the convolution in equation~\ref{eq:ReucedFormWithKernelDecomp}, which is computationally demanding, with the convolution in equation~\ref{eq:DefPsi}. This provides a dramatic increase in estimation speed, since the matrix $\boldsymbol\Psi(\mathbf{r}')$ can be calculated analytically. This is of great importance as it lowers the computational demands of the estimation algorithm.

As a result of the model reduction via the basis function decomposition and the decomposition of the connectivity kernel the state space model is formed
\begin{align}
 \mathbf x_{t+1}&=q(\mathbf x_t)\boldsymbol\theta+\xi \mathbf x_t+\boldsymbol e_t(\mathbf r) \label{eq:StateEq} \\
 \mathbf y_t&=\mathbf C \mathbf x_t+\boldsymbol \epsilon_t,\label{eq:OutputEq}
\end{align}
where the nonlinear function, $q(\cdot)$, is defined as
\begin{equation}\label{eq:QmatrixForSigmapoints}
	q(\mathbf{x}_t) = \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) d\mathbf{r}'.
\end{equation}
\section{Estimation Algorithm}\label{sec:EstimationAlgorithm}
Here we describe the problem of estimating non-linear state space neural field model using expectation-maximisation (EM) based algorithm \cite{Dempster1977,Shumway2000}. EM is an iterative algorithm that converges on the maximum-likelihood estimates of the parameters. Both the E and the M steps of the EM algorithm increase a lower bound on the log-likelihood function of the parameters given the observed field.  In the E-step we estimate the state $\mathbf x_t$, of the system given the observed field and current estimate of the model parameters to calculate expected log-likelihood. In the M-step we estimate the system parameters by maximising the expected log-likelihood computed in the E-step. The resulting parameters are then used to estimate the distribution of the state vector sequence for the next iteration. The stopping criterion is usually associated with either the change in the parameter estimates or the log-likelihood variation \cite{McLachlan1997}.
\subsection{E-Step}
In state space models, E-step corresponds to solving the smoothing problem, i.e. estimating the states of the system, $\mathbf x_t$ given all observations and the parameters of the system. The smoother output is then used to calculate the expected log-likelihood function required in M-step. The non-linear behaviour of the state-space neural field described in \eqref{eq:StateEq} requires a non-linear smoother to be employed for state estimation in the E-step.
% In order to caulculate the expected log-likelihood in the E-step the smoothed states of the system is required.  An additive form of the unscented Rauch-Tung-Striebel smoother (URTSS)~\cite{Sarkka2010} was used to estimate the approximate state distribution in the E-step.
% It can be shown that the maximum in the E-step occurs when the state distribution is exactly $p(\mathbf X\mid \mathbf Y;\boldsymbol \theta')$, the conditional distribution of X, given all observations and the current value of the parameter estimates which corresponds to solving the smoothing problem. 
\subsubsection{State estimation}
The additive form of the unscented Rauch-Tung-Striebel smoother (URTSS)~\cite{Sarkka2010} was used for the state estimation. The URTSS incorporates an unscented Kalman filter (UKF) \cite{Julier1997, Merwe2003} in a forward step to estimate posterior states, $\hat{\mathbf x}_t^{f}$, followed by a backward step to compute the smoothed state estimates, $\hat{\mathbf x}_t^{b}$. The first and second order moments of the predicted state are captured by propagating the so-called sigma points through the state equation. The sigma points, $\mathcal X_i$, are calculated using the unscented transform as follows:
\begin{align}\label{eq:sigmapoints1}
	\mathcal X_{0}&=\mathbf{\bar x} \\
	\mathcal X_{i}&= \mathbf{\bar x}+\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i, \quad i=1, \dots, n_x \\
	\mathcal X_{i}&=\mathbf{\bar x}-\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_{i- n_x}, \quad i= n_x+1, \dots, 2n_x 
\end{align}
where $\mathbf{\bar x}$ represents either $\hat{\mathbf x}_t^{f}$ or $\hat{\mathbf x}_t^{b}$, $\mathbf{P}_x$ is the corresponding covariance matrix for filtering or smoothing, $\left(\sqrt{( n_x + \lambda)\mathbf P_x}\right)_i$ is the $i^{th}$ column of the scaled matrix square root of $\mathbf P_x$, and $n_x$ is the dimension of the state space. The total number of sigma points is $2n_x+1$. The scaling parameter, $\lambda$, is defined as 
\begin{equation}\label{eq:sigmapoints3}
	\lambda=\alpha^2( n_x+\kappa) - n_x, 
\end{equation}
The positive constant $\alpha$, that determines the spread of the sigma points around $\mathbf{\bar x}$, was set to $10^{-3}$. It is typically set arbitrary small to minimise higher-order effects \cite{Haykin2001}. The other scaling parameter, $\kappa$, was set to $3-n_x$ in accordance to \cite{Julier2002a}.

The sigma vectors are propagated through the system equations and weighted to form the predicted mean and covariance. The weights are calculated by 
\begin{align}
	\mathbf W_0^{(m)}&=\frac{\lambda}{ n_x+\lambda} \\
	\mathbf W_0^{(c)}&=\frac{\lambda}{ n_x+\lambda}+(1-\alpha^2+\beta) \\
	\mathbf W_i^{(m)}&=\mathbf W_i^{(c)}=\frac{1}{2( n_x+\lambda)} \quad i=1, \dots, 2n_x, 
\end{align}
where the superscripts $m$ and $c$ denote mean and covariance and $\beta$ incorporates prior knowledge of the distribution of the states, $\mathbf{x}$ (for a Gaussian disturbance, $\beta$ should be set to 2 \cite{Haykin2001}). Since the observation equation is linear (equation~\ref{ObservationEquation}), the standard Kalman Filter update equations are used to correct the predicted states. The state estimates from forward filtering are used to form a new set of sigma points for the smoother, as described above. The predicted and the smoothed cross-covariance matrices of the states, $\mathbf M_{t}^{b-}$ and  $\mathbf M_{t}^{b}$, are also required. The former is used  for computing the smoother gain and the latter is used to compute expected log-likelihood function. The formulation of the cross-covariance matrices are given in the summary of the steps in the URTSS algorithm in Table~\ref{tab:URTSSAlgorithm}.

\subsubsection{Expected log-likelihood} 
It is well known that the expected joint data log likelihood $\ln p(\mathbf X,\mathbf Y;\boldsymbol\Theta)$, forms a lower bound on the log-likelihood function $\ln p(\mathbf Y;\boldsymbol\Theta)$, where $\mathbf X$ and $\mathbf Y$ are set of all states and the set of all observations respectively and $\boldsymbol \Theta$ denotes the set of all system parameters, i.e.
\begin{equation}
	\boldsymbol\Theta = \left\lbrace \boldsymbol\theta, \xi, \sigma_d, \sigma_{\epsilon}\right\rbrace 
\end{equation}
 Therefore we can write
\begin{align}\label{eq:QFunction}
 \mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')&=\mathbf E_{\boldsymbol \Theta'}\left[2\ln p(\mathbf X,\mathbf Y;\boldsymbol \Theta)\right],
\end{align}
where $ \mathbf E_{\boldsymbol \Theta'}\left[ .\right] $ denotes expectation taken with respect to the distribution $p(\mathbf X\mid\mathbf Y;\boldsymbol \Theta'$) which is marginal distribution of the state estimates conditioned on observed field and current estimate of the parameter set $\boldsymbol\Theta'$, i.e. smoother output. The factor of $2$ is included to simplify the derivation in the later steps. We proceed by expanding the joint distribution $p(\mathbf X,\mathbf Y;\boldsymbol \Theta)$ as 
\begin{equation}\label{eq:jointdistribution}
 p(\mathbf X,\mathbf Y;\boldsymbol \Theta)=\prod_{t=0}^{T-1} p(\mathbf y_{t+1}|\mathbf x_{t+1}; \sigma_{\epsilon})p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta, \xi, \sigma_d)p(\mathbf x_0).
\end{equation}
 Substituting equations \eqref{eq:jointdistribution} in \eqref{eq:QFunction} gives
\begin{align}
 \mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')&=\mathbf E_{\boldsymbol\Theta'}\left[\sum_{t=0}^{T-1}2\ln p(\mathbf y_{t+1}|\mathbf x_{t+1};\sigma_{\epsilon})+\sum_{t=0}^{T-1}2\ln p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta,\xi, \sigma_d)+\sum_{t=0}^{T-1}2\ln p(\mathbf x_0)\right].
\end{align}
 It should be noted that the last term is not a function of parameter set and can be dropped, therefore $\mathcal Q$-function can be rewritten as
\begin{equation}\label{eq:QIntermsofJointDist}
\mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}2\ln p(\mathbf y_{t+1}|\mathbf x_{t+1}; \sigma_{\epsilon})+\sum_{t=0}^{T-1}2\ln p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta, \xi ,\sigma_d)\right].
\end{equation}
Gaussian distributions of the state disturbance and the observation noise results into the following conditional distributions, where the normalising factors are ignored 
\begin{align}
 p\left(\mathbf y_{t+1}|\mathbf x_{t+1};\sigma_{\epsilon}\right)&= \mid\boldsymbol\Sigma_{\epsilon}\mid^{-\frac{1}{2}}  \exp\left({-\frac{1}{2}\left(\mathbf y_{t+1}-\mathbf C\mathbf  x_{t+1}\right)^\top\boldsymbol\Sigma_{\epsilon}^{-1}\left(\mathbf y_{t+1}-\mathbf C\mathbf  x_{t+1}\right)}\right),\\
p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta ,\sigma_d)&= \mid\boldsymbol\Sigma_{e}\mid^{-\frac{1}{2}} \exp \left(-\frac{1}{2}(\mathbf x_{t+1}-q(\mathbf  x_t)\boldsymbol\theta-\xi  \mathbf x_t)^\top\boldsymbol\Sigma_e^{-1}(\mathbf x_{t+1}-q( \mathbf x_t)\boldsymbol\theta-\xi \mathbf  x_t) \right)
\end{align}
Note that the disturbance covariance matrix can be written as a product of the unknown parameter, $\sigma_d^2$ and a constant matrix, $\tilde{\boldsymbol\Sigma}_e$ where the constant part is calculated using the estimated support from the correlation analysis. By expanding the exponent and taking $2\ln$ we get
\begin{align}\label{eq:CondititionDist1}
2\ln p\left(\mathbf y_{t+1}|\mathbf x_{t+1};\sigma_{\epsilon}\right)&=-n_y\ln \sigma_{\epsilon}^2-\frac{1}{\sigma_{\epsilon}^2}\left[ (\mathbf y_{t+1}-\mathbf C\mathbf  x_{t+1})^\top(\mathbf y_{t+1}-\mathbf C\mathbf  x_{t+1})\right],  \\
2\ln p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta ,\sigma_d)&=-n_x\ln\sigma_d^2-\ln\mid\tilde{\boldsymbol\Sigma}_e\mid-\frac{1}{\sigma_d^2}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}\mathbf x_{t+1}+\frac{2}{\sigma_d^2}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q( \mathbf x_t)\boldsymbol\theta+\frac{2\xi}{\sigma_d^2}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}\mathbf x_t \nonumber \\
&-\frac{1}{\sigma_e^2}\boldsymbol\theta^\top q^\top(\mathbf x_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\boldsymbol\theta-\frac{2\xi}{\sigma_d^2} \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\boldsymbol\theta-\frac{\xi^2}{\sigma_d^2}\mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}\mathbf x_t. \label{eq:CondititionDist2}
\end{align}
\dean{Dean to check expansion.} Now we rearrange the equation so we can  use the information from the smoother. (or isolate (reason for doing this step is..)) the cross-covariance etc. Taking trace and rearranging, using the invariant cyclic permutations property of the trace, equations \eqref{eq:CondititionDist1} and \eqref{eq:CondititionDist2} become
\begin{align}\label{eq:Qfunctiontrace1}
  2\ln p\left(\mathbf y_{t+1}|\mathbf x_{t+1};\sigma_{\epsilon}\right) &= -n_y\ln \sigma_{\epsilon}^2-\frac{1}{\sigma_{\epsilon}^2}\mathrm{tr}\left\lbrace(\mathbf y_{t+1}-\mathbf C\mathbf  x_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf  x_{t+1})^\top\right\rbrace    \\
 2 \ln p(\mathbf x_{t+1}|\mathbf x_{t};\boldsymbol \theta ,\sigma_d) &=-n_x\ln\sigma_d^2-\frac{1}{\sigma_d^2}\mathrm{tr}\left\lbrace\mathbf x_{t+1} \mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace+\frac{2}{\sigma_d^2}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q( \mathbf x_t)\boldsymbol\theta+\frac{2\xi}{\sigma_d^2} \mathrm{tr} \left\lbrace \mathbf x_t\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace \nonumber \\
&-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top q^\top(\mathbf x_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\boldsymbol\theta-\frac{2\xi}{\sigma_d^2} \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\boldsymbol\theta-\frac{\xi^2}{\sigma_d^2}\mathrm{tr}\left\lbrace \mathbf x_t \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace. \label{eq:Qfunctiontrace2}
\end{align}
Substituting  \eqref{eq:Qfunctiontrace1} and \eqref{eq:Qfunctiontrace2} into \eqref{eq:QIntermsofJointDist} and taking the expectation of the log likelihood function for all time instants gives the lower-bound which is to be maximised in the M-step. 
\begin{align}\label{eq:Voldermont}
 \mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')&= -(T-1)n_y\ln \sigma_{\epsilon}^2-\frac{1}{\sigma_{\epsilon}^2}\mathrm{tr}\left\lbrace\boldsymbol\sum_{t=0}^{T-1}\left[ (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top\right] \right\rbrace\nonumber \\
&-(T-1)n_x\ln\sigma_d^2-\frac{1}{\sigma_d^2}\mathrm{tr}\left\lbrace \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}\mathbf x_{t+1}^\top\right]\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace +\frac{2\xi}{\sigma_d^2} \mathrm{tr}\left\lbrace \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t+1}^\top\right] \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace \nonumber \\
&-\frac{\xi^2}{\sigma_d^2}\mathrm{tr} \left\lbrace\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t}^\top\right]\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace +\frac{2}{\sigma_d^2}\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q( \mathbf x_t)\right]\boldsymbol\theta
 \nonumber \\
&-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}  q^\top(\mathbf  x_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\right]\boldsymbol\theta-\frac{2\xi}{\sigma_d^2} \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1} \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)  \right] \boldsymbol\theta
\end{align}
The second term in \eqref{eq:Voldermont} is obtained by adding and subtracting $\mathbf C \mathbf{\hat{x}}_{t+1} \mathbf{\hat{x}}_{t+1}^\top \mathbf C^\top$ to the expectation of the second term in \eqref{eq:Qfunctiontrace1}. By defining $\Xi$-variables as 
\begin{align}
	\boldsymbol\Xi_{0}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}\mathbf x_{t+1}^\top\right]\label{eq:defofXi0} \\
\boldsymbol\Xi_{1}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t+1}^\top\right]\label{eq:defofXi1} \\
\boldsymbol\Xi_{2}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_t\mathbf x_{t}^\top\right]\\
\boldsymbol\Xi_{3}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q( \mathbf x_t)\right] \label{eq:Xi3} \\	
\boldsymbol\Xi_{4}&= \mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1}  q^\top(\mathbf  x_t)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)\right] \label{eq:Xi4}\\
 \boldsymbol\Xi_{5}&=\mathbf E_{\Theta'}\left[\sum_{t=0}^{T-1} \mathbf x_t^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf x_t)  \right] \label{eq:Xi5}
 \end{align}
equation \eqref{eq:Voldermont} can be written as
\begin{align}\label{eq:VoldermontwitXivariables}
 \mathcal Q(\boldsymbol \Theta,\boldsymbol\Theta')&=-(T-1)n_y\ln \sigma_{\epsilon}^2-\frac{1}{\sigma_{\epsilon}^2}\mathrm{tr}\left\lbrace\boldsymbol\sum_{t=0}^{T-1}\left[ (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top\right] \right\rbrace\nonumber \\
&-(T-1)n_x\ln\sigma_d^2-\frac{1}{\sigma_d^2}\mathrm{tr}\left\lbrace \boldsymbol\Xi_{0} \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace +\frac{2\xi}{\sigma_d^2} \mathrm{tr}\left\lbrace \boldsymbol\Xi_{1} \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace -  \frac{\xi^2}{\sigma_d^2}\mathrm{tr} \left\lbrace\boldsymbol\Xi_{2}\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace 
 \nonumber \\
&+\frac{2}{\sigma_d^2}\boldsymbol\Xi_{3}\boldsymbol\theta-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top \boldsymbol\Xi_{4}\boldsymbol\theta-\frac{2\xi}{\sigma_d^2}\boldsymbol \Xi_{5} \boldsymbol\theta.
\end{align}
Covariance and cross-covariances in equation \eqref{eq:VoldermontwitXivariables}, consist of both linear terms, $\boldsymbol\Xi_0$, $\boldsymbol\Xi_1$ and $\boldsymbol\Xi_2$ and non-linear terms, $\boldsymbol\Xi_3$, $\boldsymbol\Xi_4$ and  $\boldsymbol\Xi_5$. Finding linear $\boldsymbol\Xi$-variables is straightforward as they can be computed directly using smoother outputs, i.e. $\hat{\mathbf x}_t^b $, $\mathbf P_t^b$ and $\mathbf M_t^b$ using following expressions \cite{Gibsona2005} (see supporting information for details)
\begin{align}
	\boldsymbol\Xi_{0}&=\sum_{t=0}^{T-1}\left(\mathbf P_{t+1}^b+\mathbf{\hat x}_{t+1}^b\mathbf{\hat x}_{t+1}^{b\top}\right)\label{eq:defofXi0}, \\
\boldsymbol\Xi_{1}&=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}^b+\mathbf{\hat x}_t^b\mathbf{\hat x}_{t+1}^{b\top}\right) \label{eq:defofXi1}, \\
\boldsymbol\Xi_{2}&=\sum_{t=0}^{T-1}\left(\mathbf P_t^b+\mathbf{\hat x}_t^b\mathbf{\hat x}_t^{b\top}\right).
 \end{align}
However calculation of non-linear $\boldsymbol\Xi$-variables is not straightforward as in linear case. Therefore, an additional  Taylor expansion step to the URTSS algorithm is also created to approximate $q(\mathbf{x}_t)$ at each time instant which is to be used in  calculations of $\boldsymbol\Xi_3$, $\boldsymbol\Xi_4$ and $\boldsymbol\Xi_5$. This is accomplished by taking the first-order truncated Taylor-series expansion of the sigmoidal activation function  around the smoothed state, $\hat{\mathbf x}_t^b$ at each time instant, i.e.
\begin{equation}\label{eq:SigmoidTaylor}
 f\left(\boldsymbol \phi^\top\left(\mathbf r'\right)\mathbf x_t\right) \approx f\left(\boldsymbol \phi^\top(\mathbf r')\hat{\mathbf x}_t^b\right)+  \boldsymbol \phi^\top\left(\mathbf r'\right)\left(\mathbf x_t - \hat{\mathbf x}_t^b\right)f'\left(\boldsymbol \phi^\top\left(\mathbf r'\right)\hat{\mathbf x}_t^b\right)
\end{equation}
where 
\begin{equation}\label{eq:SigmoidDerivative}
 f'\left(\boldsymbol \phi^\top\left(\mathbf r'\right)\hat{\mathbf x}_t^b\right)=\varsigma f\left(\boldsymbol \phi^\top\left(\mathbf r'\right)\hat{\mathbf x}_t^b\right)\left( 1-f\left( \boldsymbol \phi^\top\left(\mathbf r'\right)\hat{\mathbf x}_t^b\right)\right)
\end{equation}
substituting \eqref{eq:SigmoidTaylor} in \eqref{eq:QmatrixForSigmapoints}  we get an approximation for $q(\mathbf x_t)$ around the smoothed state $\hat{\mathbf{x}}_t^b$ 
\begin{equation}\label{eq:qTaylor}
 q_L(\mathbf x_t) \approx q\left(\hat{\mathbf x}_t^b\right)+\int_\Omega \boldsymbol{\Psi}\left(\mathbf{r}'\right)\boldsymbol \phi^\top\left(\mathbf r'\right) \left(\mathbf x_t -\hat{\mathbf x}_t^b\right)f'\left(\boldsymbol \phi^\top\left(\mathbf r'\right)\hat{\mathbf x}_t^b\right) d\mathbf{r}'
\end{equation}
Having approximated $q\left(\mathbf x_t\right)$, by substituting  \eqref{eq:qTaylor} into equations (\ref{eq:Xi3}-\ref{eq:Xi5}) an approximate to the non-linear $\boldsymbol\Xi$-variables can be calculated (see supporting information) which in turn provides an approximate lower bound on the log-likelihood function, 
\begin{align}
\boldsymbol\Xi_{3}&\approx\sum_{t=0}^{T-1}\left[ \mathbf{\hat x}_{t+1}^\top\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t^b)+\int_\Omega\boldsymbol \phi^\top(\mathbf r') \mathbf M_{t+1}^b \tilde{\boldsymbol\Sigma}_e^{-1} \boldsymbol{\Psi}(\mathbf{r}') f'(\boldsymbol \phi^\top(\mathbf r')\mathbf {\hat x}_t^b) d\mathbf{r}'\right] \\	
\boldsymbol\Xi_{4}&\approx\sum_{t=0}^{T-1}\left[q^\top(\mathbf{\hat x}_t^b)\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t^b)+\Lambda_t\right] \\
 \boldsymbol\Xi_{5}&\approx\sum_{t=0}^{T-1}\left[ \mathbf{\hat x}_{t}^{b\top}\tilde{\boldsymbol\Sigma}_e^{-1}q(\mathbf{\hat x}_t^b)+\int_\Omega\boldsymbol \phi^\top(\mathbf r') \mathbf P_t^b \tilde{\boldsymbol\Sigma}_e^{-1}  \boldsymbol{\Psi}(\mathbf{r}') f'(\boldsymbol \phi^\top(\mathbf r')\mathbf {\hat x}_t^b) d\mathbf{r}'\right] 
 \end{align}
where
\begin{align}\label{eq:lambdat} 
	{\boldsymbol \lambda}_t^{(i)}&=\int_{\Omega} \boldsymbol \Psi(\mathbf r')\phi_i(\mathbf r')f'(\boldsymbol\phi^\top(\mathbf r')\mathbf {\hat x}_t^b)d\mathbf r'
\end{align}
and
\begin{equation}
 \boldsymbol\Lambda_t=\sum_{i,j=1}^{n_x}[\mathbf{P}_t^b]_{ij}{\boldsymbol\lambda}_t^{(i)\top}\tilde{\boldsymbol\Sigma}_e^{-1}{\boldsymbol\lambda}_t^{(j)}
\end{equation}
By computing $\boldsymbol\Xi$-variables, calculation of the lower bound is complete. The lower bound now can be used in M-step to find a set of new parameters. An approximate for computational complexity of the algorithm is given in Table~\ref{table:ComputationalComplexity}. Note that computational complexity of the nonlinear terms depend on the number of points, $n_r$ used for spatial discritisation.  In this table only the computational complexity of variables specific to the presented algorithm are shown, and the computational complexity of the URTSS can be found elsewhere \cite{Sarkka2010}.

Define
\begin{equation}
	e(x) = q(x) - q_L(x) \nonumber
\end{equation}

Then we can write $q(x) = q_L(x) + e(x)$ and use this to show that the error in the estimate is proportional to $e(x)$.

Equation \ref{eq:qTaylor} is used to approximate the lower bound on the log-likelihood function in the M-step. \dean{The difference between the linearised and nonlinear likelihood will be the error in the approximation at the state estimate that maximised the likelihood.}
\subsection{M-Step}
% In the M-step, an update $\boldsymbol\Theta'$ is chosen such that $\mathcal{Q}\left(\boldsymbol\Theta',\boldsymbol\Theta'\right)>\mathcal{Q}\left(\boldsymbol\Theta,\boldsymbol\Theta'\right)$, this will guarantee that the likelihood increases.
The objective of the M-step is to maximize the lower bound found in E-step, whcich consequenly increases the log-likelihood function $p(\mathbf Y;\boldsymbol\Theta)$.
% on the joint likelihood function, $p(\mathbf X,\mathbf Y;\boldsymbol \Theta)$, where $\mathbf X$ is the set of all states, $\mathbf Y$ is the set of all observations, and $\boldsymbol \Theta$ denotes the set of all unknown parameters, i.e., connectivity kernel parameters $\boldsymbol\theta$, synaptic dynamic $\xi$, disturbance variance $\sigma^2_d$ and the observation noise variance $\sigma^2_{\epsilon}$.
                                                                                                                                                                                                                                                                                                                                                                                                                                                                            Consider first the maximisation of $\mathcal{Q}$-function with respect to $\xi$, we can write
\begin{equation}\label{eq:Qpartialxi}
 \frac{\partial \mathcal Q}{\partial\xi}=\frac{2}{\sigma_d^2}\mathrm{tr}\left\lbrace \boldsymbol\Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace -\frac{2\xi}{\sigma_d^2}\mathrm{tr}\left\lbrace \boldsymbol\Xi_2 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-\frac{2}{\sigma_d^2}\boldsymbol\Xi_5,
\end{equation}
setting \eqref{eq:Qpartialxi} to zero and solving for $\xi$ gives
\begin{equation}
 \hat{\xi}=\frac{\mathrm{tr}\left\lbrace \boldsymbol\Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-\boldsymbol\Xi_5\boldsymbol\Xi_4^{-1}\boldsymbol\Xi_3^\top}{\mathrm{tr}\left\lbrace \boldsymbol\Xi_2 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-\boldsymbol\Xi_5\boldsymbol\Xi_4^{-1}\boldsymbol\Xi_5^{\top}}
\end{equation}
In order to maximise $\mathcal{Q}$-function with respect to $\boldsymbol\theta$ we can write
\begin{equation}\label{eq:Qpartialtheta}
 \frac{\partial \mathcal Q}{\partial \boldsymbol\theta}=\frac{2}{\sigma_d^2}\boldsymbol\Xi_3-\frac{1}{\sigma_d^2}\boldsymbol\theta^\top(\boldsymbol\Xi_4^\top+\boldsymbol\Xi_4)-\frac{2\xi}{\sigma_d^2}\boldsymbol\Xi_5,
\end{equation}
noting that variable $\boldsymbol\Xi_4$ is symmetric by construction, equating \eqref{eq:Qpartialtheta} to the zero vector yields
\begin{equation}\label{eq:theta}
 \hat{\boldsymbol \theta}=\boldsymbol\Xi_4^{-1}\left(\boldsymbol\Xi_3^\top-\xi\boldsymbol\Xi_5^\top \right).
\end{equation}
We comment on the condition under which $\boldsymbol\Xi_4$ is invertible later on. Now maximizing with respect to $\sigma_{\epsilon}^2$ we get
\begin{equation}
  \frac{\partial \mathcal Q}{\partial \sigma_{\epsilon}^2}=-(T-1)n_y\frac{1}{\sigma_{\epsilon}^2}+\frac{1}{\sigma_{\epsilon}^4}\mathrm{tr}\left\lbrace\sum_{t=0}^{T-1}\left[ (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top\right] \right\rbrace
\end{equation}
setting above to zero and solving for $\sigma_{\epsilon}^2$ gives
\begin{equation}
 \hat{\sigma}_{\epsilon}^2=\frac{1}{(T-1)n_y}\sum_{t=0}^{T-1}\mathrm{tr}\left\lbrace (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1}) (\mathbf y_{t+1}-\mathbf C\mathbf{\hat{x}}_{t+1})^\top+\mathbf C \mathbf P_{t+1}\mathbf C^\top \right\rbrace
\end{equation}
Finally differentiating with respec to $\sigma_d^2$ we get
\begin{align}
 \frac{\partial \mathcal Q}{\partial \sigma_d^2}&= -(T-1)n_x\frac{1}{\sigma_d^2}+\frac{1}{\sigma_d^4}\mathrm{tr}\left\lbrace \boldsymbol\Xi_0 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-\frac{2\xi}{\sigma_d^4} \mathrm{tr}\left\lbrace \boldsymbol\Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace+\frac{\xi^2}{\sigma_d^4}\mathrm{tr} \left\lbrace\boldsymbol\Xi_2\tilde{\boldsymbol\Sigma}_e^{-1} \right\rbrace  \nonumber \\
&-\frac{2}{\sigma_d^4}\boldsymbol\Xi_3\boldsymbol\theta +\frac{1}{\sigma_d^4}\boldsymbol\theta^\top \boldsymbol\Xi_4\boldsymbol\theta+\frac{2\xi}{\sigma_d^4} \boldsymbol\Xi_5 \boldsymbol\theta 
\end{align}
setting above to zero and solving for $\sigma_d^2$ we have
\begin{align}
\hat{\sigma}_d^2=\frac{1}{(T-1)n_x}\left[ \mathrm{tr}\left\lbrace \boldsymbol\Xi_0 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace-
2\xi\mathrm{tr}\left\lbrace \boldsymbol\Xi_1 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace +\xi^2\mathrm{tr} \left\lbrace\boldsymbol\Xi_2\tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace -2\boldsymbol\Xi_3\boldsymbol\theta+\boldsymbol\theta^\top\boldsymbol \Xi_4\boldsymbol\theta+2\xi\boldsymbol\Xi_5 \boldsymbol\theta\right]  
\end{align}
In order to confirm calculated parameters are indeed maxima of the lower bound, the second derivative test is carried out for each parameter, giving
\begin{alignat}{2}%(number of &)+1 divided by 2 here we have (3+1)/2
 \frac{\partial \mathcal Q^2}{\partial^2\xi}&\biggr\rvert_{\xi=\hat{\xi}}&&=-\frac{2}{\sigma_d^2}\mathrm{tr}\left\lbrace \boldsymbol\Xi_2 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace \label{eq:SecondDerivativexi}\\
 \frac{\partial \mathcal Q^2}{\partial^2\boldsymbol\theta}&\biggr\rvert_{\boldsymbol\theta=\hat{\boldsymbol\theta}}&&=-\frac{2}{\sigma_d^2}\boldsymbol\Xi_4 \label{eq:SecondDerivativetheta} \\
 \frac{\partial \mathcal Q^2}{\partial^2\sigma_{\epsilon}^2}&\biggr\rvert_{\sigma_{\epsilon}=\hat{\sigma}_{\epsilon}}&&=-\frac{(T-1)n_y}{\hat{\sigma}_{\epsilon}^4} \label{eq:SecondDerivativenoise} \\
 \frac{\partial \mathcal Q^2}{\partial^2\sigma_{d}^2}&\biggr\rvert_{{\sigma}_{d}=\hat{\sigma}_{d}}&&= -\frac{(T-1)n_x}{\hat{\sigma}_{d}^4} \label{eq:SecondDerivativedisturbance}
\end{alignat}
$\boldsymbol\Xi_2$ is positive definite under the persistent excitation condition of state vectors, which is guaranteed by the disturbance, $\mathbf e$, and $\tilde{\boldsymbol\Sigma}_e^{-1}$ is positive definite by construction (Note that the inverse of a positive definite matrix is also positive definite \cite{Horn2005}), therefore we have \cite{Coope1994}
\begin{equation}
 \mathrm{tr}\left\lbrace \boldsymbol\Xi_2 \tilde{\boldsymbol\Sigma}_e^{-1}\right\rbrace >0,
\end{equation}
meaning \eqref{eq:SecondDerivativexi} is negative and $\hat{\xi}$ represents a maximum. Equation \eqref{eq:SecondDerivativetheta} is negative where $\boldsymbol\Xi_4$ is positive definite requiring $\mathbf q(\mathbf x_t)$ to be of  rank $n_{\theta}$ \cite{Golub1996}. The matrix $\boldsymbol\Psi(\mathbf r')$ defined by \eqref{eq:DefPsi} is of rank $n_{\theta}$ as each row of the matrix is a set of equally spaced Gaussian basis functions with equal widths, consequently each row can be considered as a Riesz basis function for the subspaces they generate \cite{Ferrari2004}. (note that multiplying by $\mathbf \Gamma^{-1}$ doesn't change the rank). However, $q(\mathbf x_t)$ defined in \eqref{eq:QmatrixForSigmapoints} cannot be computed analytically as the inner product of a Gaussian and a sigmoid cannot be solved analytically, making it difficult to calculate the rank of $q(\mathbf x_t)$ . Therefore, we recommend to check the rank of $q(\mathbf x_t)$ at each iteration of the E-step.
% to show this we define
% \begin{equation}
%  \boldsymbol\Psi(\mathbf r')=T_s\boldsymbol\Gamma^{-1}\tilde{\boldsymbol\Psi}(\mathbf r')
% \end{equation}
% where 
% \begin{equation}\label{eq:DefPsitilde}
% 	\left[ \tilde{\boldsymbol\Psi}(\mathbf{r}')\right]_{:i} =\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\psi_i (2c_i+\mathbf{r}'-\mathbf{r})\textrm{d}\mathbf{r}},
% \end{equation}
% since matrix $\boldsymbol\Gamma^{-1}$ is symmetric
From \eqref{eq:SecondDerivativenoise} and \eqref{eq:SecondDerivativedisturbance}, observation noise and disturbance covariance estimates clearly represent maxima of the $\mathcal{Q}$-function. 
%  Increasing the joint likelihood function increases the likelihood function $p(\mathbf Y;\boldsymbol \Theta)$ \cite{McLachlan1997}. 
 \begin {table}
\begin{center}
\scalebox{1}{\begin{tabular}{lll}
\hline \hline
Variable &Order&Order of the approximate\\ \hline\\
$\Xi_0, \Xi_1, \Xi_2$&$O(Tn_x^2)$&-\\
$\Xi_3, \Xi_5$&$O(Tn_x^2n_{\theta})$&$O(Tn_x^2n_r^2)+O(Tn_xn_{\theta}n_r^2)$ \\
$\Xi_4$&$O(Tn_x^2n_{\theta})+O(Tn_xn_{\theta}^2)$&$O(Tn_x^4n_{\theta}n_r^2)+O(Tn_x^3n_{\theta}^2n_r^2)$\\
$\hat{\boldsymbol\theta} $&$O(n_{\theta}^3)$&-\\
$\hat{\xi}$&$O(n_x^3+n_{\theta}^3)$&-\\
 $\hat{\sigma}_{\epsilon}^2$&$O(Tn_y^2n_x^2)$&-\\
$\hat{\sigma}_{d}^2$&$O(Tn_x^3)+O(Tn_{\theta}^3)$&-\\
\hline \hline
\end{tabular}}
 \caption {Computational complexity of EM algorithm} 
\label{table:ComputationalComplexity}
\end{center}
\end {table}
\section{Results}
This section demonstrates the identification of the proposed IDE model using simulated data. The observation noise and disturbance were unknown to the estimation algorithm and no assumption were made on the shape of the kernel.  Data was generated employing state space equation (\ref{eq:StateEq}-\ref{eq:QmatrixForSigmapoints}) using a square grid of $n_x= 9 \times 9$ equally spaced ($2.5~mm$) basis functions with $\sigma_{\phi}^2=1.58~mm^2$, assuming the observed field can be fully described by the decomposed IDE model. 100 realizations of 500 ms of data were generated and the first 100 ms was discarded allowing the model's dynamic to reach steady state. The observation noise was set to $\Sigma_{\epsilon}=0.2\times \mathbf I_{n_y}$ and the field disturbance was set to $\mathcal{GP}(\mathbf 0,0.1 \times\gamma(\mathbf{r}-\mathbf{r}'))$ where the covariance function was chosen as a Gaussian with zero mean and the width $\sigma_{\gamma}^2=1.7~mm^2$ . The distance between adjacent sensors was 1.5 mm resulting into $n_y = 196$ observations. The width of the sensor kernels were set to $\sigma_m^2=0.81~mm^2$. The sampling period and the synaptic time constant were set to $T_s = 1~ms$  and $\tau = 10~ms$, respectively. The sigmoidal activation parameters were $\varsigma=0.56~mV^{-1}$ and $v_0=1.8~mV$.

The correlation analysis described in Section \ref{subsec:CorrAnal} were applied to generated data in order to approximate the unknown supports of both connectivity kernel and the disturbance covariance function. In the algorithm, the unknown variables $\sigma_{\epsilon}^2$ and $\xi$ in equations () and () were set to zero. By doing so, the exact shape of the kernel and disturbance covariance function cannot be inferred, this step might be considered as an initialisation for the EM algorithm, however, an approximate of their support can be still extracted. In fact, the estimated support for the disturbance covariance function was very accurate. The estimates of the connectivity kernel and disturbance support were normalised and thresholded for values less than 0.01. The result of the correlation analysis for the connectivity kernel is shown in Figure.\ref{fig:KernelWidthEstimation}. The red circle shows the estimated support which encompasses the non-zero values (black area) after thresholding. The estimated support of the disturbance covariance function convolved twice with the observation kernel is shown in Figure.\ref{fig:DisturbanceWidthEstimation}. By fitting a Guassian basis function to the result, and under the assumption that the width of the sensor,$\sigma_m^2$ were known, the width of the disturbance covariance function, $\sigma_d^2$ was calculated analytically using the expression for Gaussian basis functions convolution.

The estimated supports were then applied in the EM algorithm described in Section \ref{sec:EstimationAlgorithm} to estimate the smoothed states $\hat{\mathbf x}_t^b$, the connectivity kernel weights $\boldsymbol\theta$, the synaptic dynamic $\xi$, and the hyperparameters $\sigma_{\epsilon}^2$ and $\sigma^2_d$. A  square grid of $n_{\theta}=25$ (5 by 5) basis functions with $2.25~mm$ spacing were used to cover the estimated support where $\sigma_{\psi}^2=3.5$. This is shown in Figure.\ref{fig:KernelWidthEstimation} by blue circles, each of which representing full width at half maximum of the connectivity kernel basis functions. The true and estimated connectivity kernel for 100 Monte Carlo simulations are shown in Figure~\ref{fig:KernelEstimates}. Cross-sections of the connectivity kernel estimates for each run are shown in Figure~\ref{fig:KernelEstimateCrossSection}. The reconstructed kernel is in good accordance with the actual kernel, where the actual kernel is inside the confidence interval. The large standard deviation is due to the high number of parameters need to be estimated. Figure~\ref{fig:Histogram} depicts the histogram of parameter estimates from 100 simulations where the true parameter value is shown by the solid line and the mean parameter estimate is shown by the dashed line. These histograms indicate small bias in parameter estimates with a very small standard deviation. Examples of neural field reconstruction are given in Figure.~\ref{fig:FieldEstimates}.  The figure shows snapshots at three time instants of the true neural field and the estimated field. In each subplot neural activity is plotted against spatial index in 1D to provide a comparision between true and the estimated underlying fields. Plot of lower bound of the log likelihood versus iterations of the EM is shown in Figure.~\ref{fig:LowerBound}. In general, for the non-linear dynamical systems the actual lower bound can not be claculated analytically, therefore the approximate of the lower bound obtained by URTSS is shown. The change in lower bound drops below 0.009 percent after 15 iterations, confirming
that the algorithm converged to steady parameter values.

\begin{itemize}
	\item Will show parameter distributions
	\item Will show lower bound
      \end{itemize}

\section{Discussion}

\begin{itemize}
	\item comment on how the e step is UKF and M-step is kinda like the EKF, and we dont use EKF in E-step since UKF gives better estimates and the linearisation in M-step would be more accurate. 
	\item comment on EM has been used with EKS by Ghahramani which has poorer convergance property than using EM with URTSS
	\item itterative EM and correlation analysis (maybe)
	\item{comment on EM might convrge to suboptimal solution,simulated
annealing can be used to tackle such situations but difficult to implement (Mc Lachlan)}
	\item{since the non-linearity is a sigmoid function first-order Taylor provide a good approximation however in general settings with more complicated nonlinear behaviours higher order terms might need to be included, this however introduce intractable computational complexity.}
\end{itemize}

% % all figures here
% % ~~~~~~~~~~~~~~~~

\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/KernelBasis1.pdf}
\end{center}
\caption{{\bf Connectivity kernel support approximation.} The approximate support of the kernel is shown by the red circle. Black area shows the non-zero values of the support after thresholding. A $5\times5$ grid of basis functions (blue circles) used to reconstruct the connectivity kernel.}
\label{fig:KernelWidthEstimation}
\end{figure}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/DisturbanceWidthEstimation.pdf}
\end{center}
\caption{{\bf Disturbance support approximation}. }
\label{fig:DisturbanceWidthEstimation}
\end{figure}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/KernelEstimate.pdf}
\end{center}
\caption{{\bf Connectivity kernel estimates.} (A) Actual kernel. (B) The mean kernel estimate over 100 realisations}
\label{fig:KernelEstimates}
\end{figure}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/KernelEstimateCrossSection.pdf}
\end{center}
\caption{{\bf Cross-sections of connectivity kernel estimates.} Actual kernel and the mean kernel estimates are shown with blue and green lines respectively. Each of the estimate from Monte Carlo simulation is shown with black line. The 95\% confidence intervals are shown by shaded red area.}
\label{fig:KernelEstimateCrossSection}
\end{figure}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/Histogram.pdf}
\end{center}
\caption{{\bf Histograms of the parameter estimates over 100 realizations}. The actual and mean of the estimated parameter values are shown by solid and dotted lines respectively. (A) The histogram of the synaptic dynamic parameter, $\xi$. (B) The histogram of the observation noise variance, $\sigma_{\epsilon}$. (C) The histogram of the disturbance noise variance, $\sigma_d^2$}.
\label{fig:Histogram}
\end{figure}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\begin{figure}[!ht]
\begin{center}
\includegraphics[scale=1]{./Figures/FieldEstimates.pdf}
\end{center}
\caption{{\bf Neural field estimates at different time instants}. In each subplot, true (gray line) and reconstructed neural field (red dots) are plotted versus spatial index.}
\label{fig:FieldEstimates}
\end{figure}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\begin{figure}[!ht]
\begin{center}
\includegraphics{./Figures/Q.pdf}
\end{center}
\caption{{\bf Lower-bound of the log-likelihood}. $Q(\boldsymbol\Theta',\boldsymbol\Theta')=-(T-1)\left[n_y(1+\ln \hat{\sigma}^2_{\epsilon})+ n_x(1+\ln \hat{\sigma}^2_d)\right]$.}
\label{fig:LowerBound}
\end{figure}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/KernelWidthEstimation.pdf}
% \end{center}
% \caption{{\bf Connectivity kernel support estimation}.}
% \label{fig:KernelWidthEstimation1}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/KernelBasis.pdf}
% \end{center}
% \caption{{\bf Connectivity kernel basis functions, $n_{\theta}=25$}, $\sigma^2_{\theta}=3.5$.}
% \label{fig:KernelBasis}
% \end{figure}

% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/LFPs.eps}
% \end{center}
% \caption{{\bf Local field potentials from the micro-electrode array}. Data recorded from the first 8 channels of the array. The red dotted line indicates the seizure onset and the blue dashed line indicates the seizure end.}
% \label{fig:TimeSeries}
% \end{figure}
% 
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/FieldObservations.eps}
% \end{center}
% \caption{{\bf Snap-shots of the spatial aspects of the observed field}. The snap-shots are ordered from A to I forming a consecutive sequence spaced 2.5~ms apart starting at 34.024s}
% \label{fig:FieldObserations}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/TemporalFreq.eps}
% \end{center}
% \caption{{\bf Spectra from local field potential recordings}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
% \label{fig:TemporalFreqObservation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/SpatialFreq.eps}
% \end{center}
% \caption{{\bf Spatial frequency analysis of the observed neural field}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
% \label{fig:SpatialFreqObservation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/SpatialFreqCrossSection.eps}
% \end{center}
% \caption{{\bf Diagonal cross-section of the spatial frequency plots of the observed neural field}. A. Pre-seizure observations. B. Seizure observations. C. Post-seizure observations.}
% \label{fig:DiagSpatialFreqObservation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/CrossCorr2D.eps}
% \end{center}
% \caption{{\bf Mean cross-correlations between consecutive observations of the neural field}. Note: The correlations are normalised such that the maximum value is one. A. Pre-seizure observation correlations. B. Seizure observation correlations. C. Post-seizure observation correlations.}
% \label{fig:SpatialCrossCorrelation}
% \end{figure}
% 
% \begin{figure}[!ht]
% \begin{center}
% \includegraphics{./Figures/HomoTestCrossCorr.eps}
% \end{center}
% \caption{{\bf Mean cross-correlations between consecutive samples using subsets of observations of the neural field}. Each subplot shows correlations corresponding to a subset electrodes. Indexing the electrodes in a (x,y) grid, where index ranges from 1 to 10. Subplot A shows the mean correlation between electrode indexes (1-6, 5-10), subplot B shows correlations from indexes (5-10,5-10), subplot C shows indexes (1-6, 1-6) and subplot D shows indexes (5-10, 5-10).}
% \label{fig:HomogeneityTest}
% \end{figure}

% tables here
% ~~~~~~~~~~~
\renewcommand{\arraystretch}{1.7}
\begin{table}[!ht]
\caption{
\bf{Algorithm for the Unscented RTS Smoother}}
\label{tab:URTSSAlgorithm}
\begin{tabular}{|c|}\hline
\multicolumn{1}{|p{16cm}|}{\textbf{1.} Forward initialisation} \\ 
$\hat{\mathbf x}_0, \mathbf P_0$ \\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{2.} Forward iteration: for $t \in \left\lbrace 0,\cdots, T\right\rbrace $, calculate the sigma points $\mathcal X_{i,t}^f$ using equations \ref{eq:sigmapoints1}-\ref{eq:sigmapoints3} and propagate through equation~\ref{eq:QmatrixForSigmapoints}} \\
$\mathcal X_{i,t+1}^{f-}=Q(\mathcal X_{i,t}^f) \quad i=0, \dots, 2L$\\
\multicolumn{1}{|p{16cm}|}{Calculate the predicted state and the predicted covariance matrix} \\
$\hat{\mathbf x}_{t+1}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1}^{f-} $ \\
$\mathbf P_{t +1}^{f-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e$ \\ 
\multicolumn{1}{|p{16cm}|}{Compute the filter gain, the filtered state and the filtered covariance matrix using the standard Kalman filter update equations} \\
$\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_{\varepsilon})^{-1}$\\
$\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-}) $\\
$\mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}$\\ 
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{3.} Backward initialisation}\\
$\mathbf P_T^b= \mathbf P_T^f, \quad \hat{\mathbf x}^b_T= \hat{\mathbf x}^f_T$ \\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{4.} Backward iteration: for $t \in \left\lbrace T-1, \cdots, 0 \right\rbrace $ calculate the sigma points $\mathcal X_{i,t}^b$ and propagate them through equation \ref{eq:QmatrixForSigmapoints}}\\
$\mathcal X_{i,t+1}^{b-}=Q(\mathcal X_{i,t}^b) \quad i=0, \dots, 2L$\\
\multicolumn{1}{|p{16cm}|}{Calculate the predicted state and the predicted covariance matrix}\\
$\hat{\mathbf x}_{t+1}^{b-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1}^{b-}$\\
$\mathbf P_{t +1}^{b-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{b-})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t +1}^{b-})^\top+\boldsymbol \Sigma_e $\\
$\mathbf M_{t +1}^{b-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t}^{b-}-\hat{\mathbf x}_{t}^{f})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t+1}^{b-})^\top$\\
\multicolumn{1}{|p{16cm}|}{Compute the smoother gain, the smoothed state and the smoothed covariance matrix}\\
$\mathbf S_t=\mathbf M_{t +1}^{b-}\left[ \mathbf P_{t +1}^{b-}\right] ^{-1} $\\
$\hat{\mathbf x}_t^b=\hat{\mathbf x}_t^f+\mathbf S_t\left[\hat{\mathbf x}_{t+1}^{b}-\hat{\mathbf x}_{t+1}^{b-}\right]$\\
$\mathbf P_{t}^{b}=\mathbf P_{t}^{f}+\mathbf S_t\left[\mathbf P_{t+1}^{b}-\mathbf P_{t+1}^{b-} \right]\mathbf S_t^\top $\\
\multicolumn{1}{|p{16cm}|}{Additional step to calculate the smoothed cross covariance matrix \cite{Sarkka2008}} \\
$\mathbf M_{t +1}^{b}=\mathbf S_t\mathbf P_{t+1}^{b}$\\
\hline
\end{tabular}
\begin{flushleft}This table shows the steps in the unscented Rauch-Tung-Striebel smoother algorithm. The steps are iterated 11 times for our state estimation procedure. The least squares algorithm is run after each iteration to update the parameter estimates.
\end{flushleft}
\label{tab:UKFAlgorithm}
\end{table}
 \renewcommand{\arraystretch}{1}
\clearpage
\newpage
\bibliographystyle{plain}
\bibliography{EMIDE}
\end{document}